---
title: "S02"
author: "zifengxu"
date: "2024-05-03"
output: html_document
---


# Read the Data (PSI)
```{r}
library(haven)
library(dplyr)
library(stringr)
library(tidyr)
library(jsonlite)
```

```{r}
df <- read_sav("../../data/grade5/ASLNORZ7_processdata_PSIstudents.sav")

# Explore the data
head(df)
```

Remove the rows of auto check for this question
```{r}
df_filter = df%>%
  filter(df$PSIVariableIdentifier != 11117 |is.na(df$PSIVariableIdentifier))
```

Filter the rows and exactly one row above the row that satisfied one of the conditions below: 

Condition 1: BlockName=SQ02, BookletPart=1, BlockInPart=1, currentindex=2.

Condition 2: BlockName=SQ02, BookletPart=2, BlockInPart=2, currentindex=16.

```{r}
condition1 <- df_filter$BlockName == 'SQ02' & 
              df_filter$BookletPart == 1 & 
              df_filter$BlockInPart == 1 & 
              df_filter$currentindex == 2 

# Define the second condition
condition2 <- df_filter$BlockName == 'SQ02' & 
              df_filter$BookletPart == 2 & 
              df_filter$BlockInPart == 2 & 
              df_filter$currentindex == 16 

# Combine the two conditions using OR (|) as you want rows satisfying either condition
combined_condition <- condition1 | condition2

# Use filter to get the rows that satisfy the conditions and one row below
result2 <- df_filter %>% 
  filter(combined_condition | lag(combined_condition, default = FALSE))

result2
```


### Module Path
Create a new column module_path based on (BlockName, BookletPart, BlockInPart, currentindex)
```{r}
result_use_Q2= transform(result2,
                       module_path = paste0("(", BlockName, ", ", BookletPart, ", ", BlockInPart, ")"))
head(result_use_Q2)
```

```{r}
unique(result_use_Q2$eventname)
```


### Time
```{r}
result_filter_Q2=result_use_Q2 %>%filter(eventname %in% c("Nav:GoTo", "Response", "UI:Scrollbar", "Btn:Ruler"))
result_filter_Q2
```

```{r}
result_time_complete_Q2=result_filter_Q2 %>%
  mutate(SQ02S02_time = ifelse(currentindex == 2 & eventname == 'Nav:GoTo' | currentindex == 16 & eventname == 'Nav:GoTo', 0, ifelse(timemilisec - lag(timemilisec) > 0,
                                      (timeunixsec - lag(timeunixsec) + (timemilisec - lag(timemilisec))/1000),
                                      (timeunixsec - lag(timeunixsec) - 1 + (timemilisec + 1000 - lag(timemilisec))/1000))))
# head(result_time_complete_Q2[c("idstud","module_path","SQ02S02_time")])

# result_time_complete_Q2
```

```{r}
result_time_complete_Q2 = result_time_complete_Q2%>%
  mutate(RowNumber = row_number())
```

```{r}
result_time_complete_Q2$SQ02S02_unit_time <- result_time_complete_Q2$timeunixsec * 1000 + result_time_complete_Q2$timemilisec

# show all the units
result_time_complete_Q2$SQ02S02_unit_time <- format(result_time_complete_Q2$SQ02S02_unit_time, scientific = FALSE)
```

```{r}
result_time_complete_Q2
```

```{r}
result_time_complete_Q2=result_time_complete_Q2%>%
  mutate(information_use = case_when(
    PSIVariableIdentifier == 11112 ~ "S1",
    PSIVariableIdentifier == 11113 ~ "S2",
    PSIVariableIdentifier == 11114 ~ "S3",
    TRUE ~ information
  ))
```


### Fill in Clicking
If the column PSIVariableIdentifier == 11112, if information == 1, then Click_S1.
If the column PSIVariableIdentifier == 11113, if information == 1, then Click_S2.
If the column PSIVariableIdentifier == 11114, if information == 1, then Click_S3.

```{r}
result_click_Q2 <- result_time_complete_Q2 %>%
  mutate(information = case_when(
    PSIVariableIdentifier == 11112 & information == 1 ~ "Click_S1",
    PSIVariableIdentifier == 11113 & information == 1 ~ "Click_S2",
    PSIVariableIdentifier == 11114 & information == 1 ~ "Click_S3",
    TRUE ~ information
  ))
result_click_Q2
```


### Click in Sequence
Filter the data by PSIVariableIdentifier == 11112, 11113, 11114. Then group by the student id, the new column SQ02S02_click_sequence should created based on the sequence of the value of the column information in the order of if the row number is higher, their sequence should be higher.
```{r}
# Filter data and arrange by idstud and information
df_filtered <- result_click_Q2 %>%
  filter(PSIVariableIdentifier %in% c(11112, 11113, 11114))%>%
  group_by(idstud) %>%
  mutate(SQ02S02_sequence_list = list(information[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ02S02_click_sequence <- sapply(df_filtered$SQ02S02_sequence_list, function(x) paste0("(", toString(x), ")"))

# df_filtered
```

Create a new dataframe by merge the two dataframes by using the column idstud, and RowNumber.
```{r}
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ02S02_click_sequence")]
```

```{r}
result_click_seq_Q2 <- merge(result_click_Q2, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)
# result_click_seq_Q2
```

Then fill in all the missing value of the column SQ02S02_click_sequence according to the student name.
```{r}
result_click_seq_Q2=result_click_seq_Q2%>%
  group_by(idstud) %>%
  fill(SQ02S02_click_sequence, .direction = "downup")
result_click_seq_Q2
```


### Choice in Sequence
Filter the data by PSIVariableIdentifier == 11110. Then group by the student id, the new column SQ02S02_choice_sequence should created based on the sequence of the value of the column information in the order of if the row number is higher, their sequence should be higher.
```{r}
# Filter data and arrange by idstud and information
df_filtered <- result_time_complete_Q2 %>%
  filter(PSIVariableIdentifier == 11110)%>%
  group_by(idstud) %>%
  mutate(SQ02S02_sequence_list = list(information[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ02S02_choice_sequence <- sapply(df_filtered$SQ02S02_sequence_list, function(x) paste0("(", toString(x), ")"))

# df_filtered
```

Create a new dataframe by merge the two dataframes by using the column idstud, and RowNumber.
```{r}
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ02S02_choice_sequence")]
```

```{r}
result_choice_seq_Q2 <- merge(result_click_seq_Q2, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)
# result_choice_seq_Q2
```

Then fill in all the missing value of the column SQ02S02_choice_sequence according to the student name.
```{r}
result_choice_seq_Q2=result_choice_seq_Q2%>%
  group_by(idstud) %>%
  fill(SQ02S02_choice_sequence, .direction = "downup")
result_choice_seq_Q2
```


### Final Choice
Create a new column SQ02S02_final_choice. Filter the data by using PSIVariableIdentifier == 11110 and IsfinalAnswer ==1, then groupby idstud, SQ02S02_final_choice = information.
```{r}
# Filter the data based on PSIVariableIdentifier == 11110
filtered_data <- result_time_complete_Q2 %>% filter(PSIVariableIdentifier == 11110)%>%
filter(IsFinalAnswer == 1)

# Group by idstud and create the new column SQ02S02_final_choice
result <- filtered_data %>%
  group_by(idstud) %>%
  summarise(SQ02S02_final_choice = information)

# Merge the result back to the original data frame
result_Final_choice_Q2 <- merge(result_choice_seq_Q2, result, by = "idstud", all.x = TRUE)
```


```{r}
result_Final_choice_Q2
```


### Word Count
Filter the data based on PSIVariableIdentifier == 11111 & IsFinalAnswer == 1.
```{r}
# Filter the data based on PSIVariableIdentifier == 11111
filtered_data <- result_time_complete_Q2 %>% filter(PSIVariableIdentifier == 11111)%>%
filter(IsFinalAnswer == 1)
```

```{r}
length(unique(filtered_data$idstud))
```

Create a new column SQ02S02_wordcount_punc that count the number of words in the column information including the punctuations.
```{r}
# Function to split words and punctuations
split_words_and_punctuations <- function(text) {
  # Match words and punctuations separately
  tokens <- strsplit(text, "(?<=[a-zA-Z0-9])(?=[[:punct:]])|(?<=[[:punct:]])(?=[a-zA-Z0-9])|\\s+", perl=TRUE)[[1]]
  # Remove empty tokens
  tokens <- tokens[tokens != ""]
  return(tokens)
}

# Apply the function to the information column
filtered_data$SQ02S02_wordcount_punc <- sapply(filtered_data$information, function(x) length(split_words_and_punctuations(x)))
```
Use (?<=[a-zA-Z0-9])(?=[[:punct:]])|(?<=[[:punct:]])(?=[a-zA-Z0-9]) split words and punctuations separately, and use |\\s+ matches one or more whitespace characters. Then count the total number of tokens.

excluding the punctuations
```{r}
filtered_data <- mutate(filtered_data, SQ02S02_wordcount_no_punc = str_count(gsub("[-[:punct:]]", " ", information), "\\S+"))

filtered_data[c("idstud","information","module_path","SQ02S02_wordcount_punc","SQ02S02_wordcount_no_punc")]
```

Merge the data to the original data
```{r}
df_filtered_merge = filtered_data[c("idstud","SQ02S02_wordcount_punc","SQ02S02_wordcount_no_punc")]
```

```{r}
# Merge the result back to the original data frame
result_wordcount_Q2 <- merge(result_Final_choice_Q2, df_filtered_merge, by = "idstud", all.x = TRUE)
result_wordcount_Q2
```


### Raw Response
Create a new column SQ02S02_raw_response that is exactly same as the column information of the filtered data.
```{r}
filtered_data$SQ02S02_raw_response <- filtered_data$information

df_filtered_merge = filtered_data[c("idstud","SQ02S02_raw_response")]
```

```{r}
# Merge the result back to the original data frame
result_wordres_Q2 <- merge(result_wordcount_Q2, df_filtered_merge, by = "idstud", all.x = TRUE)
result_wordres_Q2
```


### All Steps
Create a new dataframe by copy the original one.Create a new column step_str based on the other column.
If PSIVariableIdentifier == 11112, 11113, 11114, then step_str == value of the column information
If PSIVariableIdentifier == 11110, then step_str == "select_option_" + value of the column information
If PSIVariableIdentifier == 11111, then step_str == "write_explanation"
If eventname == "Nav:GoTo" & new_currentindex == 2, then step_str == "Navigate_to_Page03_Q2"
If eventname == "Nav:GoTo" & new_currentindex != 2, then step_str == if new_currentindex in 1,2,3,5,6,7,8,9,10, then paste0("Navigate_to_Page", new_currentindex+1, "Q",cor_rule(new_currentindex)), else paste0("Navigate_to_Page", new_currentindex+1)
If eventname == "UI:Scrollbar", step_str == extract the value after event and scrollTop in the column information.
If eventname == "Btn:Ruler", step_str == "Ruler" + extract the value after To in the column information.

Create a column new_currentindex by changing the column currentindex based on the rule below:
14 to 0, 15 to 1, 16 to 2, 17 to 3, 18 to 4, 19 to 5, 20 to 6, 21 to 7, 22 to 8, 23 to 9, 24 to 10, 25 to 11.
```{r}
result_copy_Q2=data.frame(result_click_Q2)
result_copy_Q2 <- result_copy_Q2 %>%
  mutate(new_currentindex = case_when(
    currentindex == 14 ~ 0,
    currentindex == 15 ~ 1,
    currentindex == 16 ~ 2,
    currentindex == 17 ~ 3,
    currentindex == 18 ~ 4,
    currentindex == 19 ~ 5,
    currentindex == 20 ~ 6,
    currentindex == 21 ~ 7,
    currentindex == 22 ~ 8,
    currentindex == 23 ~ 9,
    currentindex == 24 ~ 10,
    currentindex == 25 ~ 11,
    TRUE ~ currentindex  # Keep the original value if it doesn't match any rule
  ))
```

create a cor_rule(new_currentindex) that:
if new_currentindex == 1,2,3,5,7,9,10,11,12, then
if new_currentindex == 1, return 1,
ifelse new_currentindex == 2, return 2,
ifelse new_currentindex == 3, return 3,
ifelse new_currentindex == 5, return 4,
ifelse new_currentindex == 6, return 5,
ifelse new_currentindex == 7, return 6,
ifelse new_currentindex == 8, return 7,
ifelse new_currentindex == 9, return 8,
ifelse new_currentindex == 10, return 9.
```{r}
cor_rule <- function(new_currentindex) {
  result <- case_when(
    new_currentindex == 1 ~ 1,
    new_currentindex == 2 ~ 2,
    new_currentindex == 3 ~ 3,
    new_currentindex == 5 ~ 4,
    new_currentindex == 6 ~ 5,
    new_currentindex == 7 ~ 6,
    new_currentindex == 8 ~ 7,
    new_currentindex == 9 ~ 8,
    new_currentindex == 10 ~ 9,
    TRUE ~ NA_integer_
  )
  return(result)
}
```

```{r}
result_str_Q2<- result_copy_Q2 %>%
  mutate(
    step_str = case_when(
      PSIVariableIdentifier %in% c(11112, 11113, 11114) ~ information,
      PSIVariableIdentifier == 11110 ~ paste0("select_option_", information),
      PSIVariableIdentifier == 11111 ~ "write_explanation",
      eventname == "Nav:GoTo" & new_currentindex == 2 ~ "Navigate_to_Page03_Q2",
      eventname == "Nav:GoTo" & new_currentindex != 2 ~ {
        case_when(
          new_currentindex %in% c(1,2,3,5,6,7,8,9,10) ~ paste0("Navigate_to_Page", new_currentindex+1, "_Q", cor_rule(new_currentindex)),
          TRUE ~ paste0("Navigate_to_Page", new_currentindex+1)
        )
      },
      eventname == "UI:Scrollbar" ~ sub('.*"event":"([^"]+).*"scrollTop":(\\d+).*', '\\1 \\2', information),
      eventname == "Btn:Ruler" ~ paste0("Ruler_", gsub('.*To":"(\\w+).*', '\\1', information)),
      TRUE ~ NA_character_  # If none of the conditions are met, set to NA
    )
  )
result_str_Q2[c("RowNumber","step_str","information")]
```
 
If the row of value of the column eventname == "UI:Scrollbar", then the value of the row of the column step_str substitute the white space to _.
```{r}
result_str_use_Q2 <- result_str_Q2 %>%
  mutate(step_str = ifelse(eventname == "UI:Scrollbar", gsub(" ", "_", step_str), step_str))
result_str_use_Q2[c("RowNumber","step_str","information")]
```

```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q2 %>%
  filter(!is.na(step_str)) %>%
  group_by(idstud) %>%
  mutate(SQ02S02_allsteps_list = list(step_str[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ02S02_allsteps_seq <- sapply(df_filtered$SQ02S02_allsteps_list, function(x) paste0("(", toString(x), ")"))
```

```{r}
df_filtered <- select(df_filtered, -SQ02S02_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ02S02_allsteps_seq")]

result_allsteps_Q2 <- merge(result_wordres_Q2, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_allsteps_Q2=result_allsteps_Q2%>%
  group_by(idstud) %>%
  fill(SQ02S02_allsteps_seq, .direction = "downup")

result_allsteps_Q2
```

```{r}
result_allsteps_Q2[c("idstud","SQ02S02_allsteps_seq","information")]
```


### Time Taken for All Steps
create a column to record the time taken for each single steps for each student to do this problem.
```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q2 %>%
  group_by(idstud) %>%
  mutate(SQ02S02_allsteps_list = list(SQ02S02_time[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ02S02_allsteps_seq_time <- sapply(df_filtered$SQ02S02_allsteps_list, function(x) paste0("(", toString(x), ")"))

# filter the data and select column to merge
df_filtered <- select(df_filtered, -SQ02S02_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ02S02_allsteps_seq_time")]

# merge the data
result_alls_time_Q2 <- merge(result_allsteps_Q2, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_alls_time_Q2=result_alls_time_Q2%>%
  group_by(idstud) %>%
  fill(SQ02S02_allsteps_seq_time, .direction = "downup")
result_alls_time_Q2
```


### Timestamp for All Steps in milisecond
create a column to record the time taken for each single steps for each student to do this problem.
```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q2 %>%
  group_by(idstud) %>%
  mutate(SQ02S02_allsteps_list = list(SQ02S02_unit_time[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ02S02_allsteps_timestamp <- sapply(df_filtered$SQ02S02_allsteps_list, function(x) paste0("(", toString(x), ")"))

# filter the data and select column to merge
df_filtered <- select(df_filtered, -SQ02S02_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ02S02_allsteps_timestamp")]

# merge the data
result_all_unit_time_Q2 <- merge(result_alls_time_Q2, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_all_unit_time_Q2=result_all_unit_time_Q2%>%
  group_by(idstud) %>%
  fill(SQ02S02_allsteps_timestamp, .direction = "downup")
result_all_unit_time_Q2
```

### Total time taken
Create a new column SQ02S02_total_time. Group by idstud, SQ02S02_total_time = sum the value of the column SQ02S02_time for each idstud.

```{r}
# Group by idstud and calculate the sum of SQ02S02_time
result <- result_time_complete_Q2 %>%
  group_by(idstud) %>%
  summarise(SQ02S02_total_time = sum(SQ02S02_time, na.rm = TRUE))

# Merge the result back to the original data frame
result_total_time_Q2 <- merge(result_all_unit_time_Q2, result, by = "idstud", all.x = TRUE)
```

```{r}
result_total_time_Q2
# result_total_time_Q2[c("idstud","SQ02S02_time","SQ02S02_total_time")]
```


### Save the data
Filter the columns of data frame, only keep the new created columns and idstud, idstud_ori, ImportLogID, BlockName, BookletPart, BlockInPart.
```{r}
resultS02 <- result_total_time_Q2 %>%
  select(idstud, idstud_ori, ImportLogID, BlockName, BookletPart,BlockInPart, module_path, SQ02S02_click_sequence, SQ02S02_choice_sequence, SQ02S02_wordcount_punc, SQ02S02_wordcount_no_punc, SQ02S02_raw_response, SQ02S02_allsteps_seq, SQ02S02_allsteps_seq_time, SQ02S02_allsteps_timestamp, SQ02S02_total_time)
# resultS02
```


```{r}
resultS02_save = resultS02[!duplicated(resultS02), ]

condition1 <- resultS02_save$BlockName == 'SQ02' & 
              resultS02_save$BookletPart == 1 & 
              resultS02_save$BlockInPart == 1 

condition2 <- resultS02_save$BlockName == 'SQ02' & 
              resultS02_save$BookletPart == 2 & 
              resultS02_save$BlockInPart == 2

combined_condition <- condition1 | condition2

resultS02_save=resultS02_save%>%ungroup()%>% 
  filter(combined_condition)
resultS02_save
```

Save the dataframe
```{r}
# write_sav(resultS02_save, "../../../new_data/grade5/Farm/SQ02S02_PSI.sav")

# write.csv(resultS02_save, file = 'csv_SQ02S02_PSI.csv', row.names = FALSE)
```






