---
title: "F01"
author: "zifengxu"
date: "2024-04-11"
output: html_document
---

SQ01,1,2,13
SQ01,2,1,1

对应关系改一下，有的页数可能不包含。

包括all steps里怎么对应

也就是说，（1）navigate的那个这么描述。（2）column name该怎么写

然后问一下，就是dataset里面，他就是其他页数也包含进去了，然后是只看题的，还是要考虑其他的页数。


尺子视频

# Read the Data (PSI)
```{r}
library(haven)
library(dplyr)
library(stringr)
library(tidyr)
library(jsonlite)
```

```{r}
df <- read_sav("../../../data/grade5/ASLNORZ7_processdata_PSIstudents.sav")

# Explore the data
head(df)
```

Remove the rows of auto check for this question
```{r}
df_filter = df%>%
  filter(df$PSIVariableIdentifier != 1003 |is.na(df$PSIVariableIdentifier))
```

First of all, filter the rows and exactly one row above the row that satisfied one of the conditions below: 

Condition 1: BlockName=SQ01, BookletPart=1, BlockInPart=2, currentindex=13.

Condition 2: BlockName=SQ01, BookletPart=2, BlockInPart=1, currentindex=1.
```{r}
condition1 <- df_filter$BlockName == 'SQ01' & 
              df_filter$BookletPart == 1 & 
              df_filter$BlockInPart == 2 & 
              df_filter$currentindex == 13 

# Define the second condition
condition2 <- df_filter$BlockName == 'SQ01' & 
              df_filter$BookletPart == 2 & 
              df_filter$BlockInPart == 1 & 
              df_filter$currentindex == 1 

# Combine the two conditions using OR (|) as you want rows satisfying either condition
combined_condition <- condition1 | condition2

# Use filter to get the rows that satisfy the conditions and one row above
result1 <- df_filter %>% 
  filter(combined_condition | lag(combined_condition, default = FALSE))
```

```{r}
result1
```


### Module Path
Create a new column module_path based on (BlockName, BookletPart, BlockInPart, currentindex)
```{r}
result_use_Q1= transform(result1,
                       module_path = paste0("(", BlockName, ", ", BookletPart, ", ", BlockInPart, ")"))
result_use_Q1
```


### Time Taken
Filter the data that only related to students actions.
```{r}
result_filter_Q1=result_use_Q1 %>%filter(eventname %in% c("Nav:GoTo", "Response", "UI:Scrollbar", "Btn:Calc"))
result_filter_Q1
```

```{r}
result_time_complete_Q1=result_filter_Q1 %>%
  mutate(SQ01F01_time = ifelse(currentindex == 1 & eventname == 'Nav:GoTo' | currentindex == 13 & eventname == 'Nav:GoTo', 0, ifelse(timemilisec - lag(timemilisec) > 0,
                                      (timeunixsec - lag(timeunixsec) + (timemilisec - lag(timemilisec))/1000),
                                      (timeunixsec - lag(timeunixsec) - 1 + (timemilisec + 1000 - lag(timemilisec))/1000))))
# head(result_time_complete_Q1[c("idstud","module_path","SQ01F01_time")])

# result_time_complete_Q1
```

```{r}
result_time_complete_Q1 = result_time_complete_Q1%>%
  mutate(RowNumber = row_number())
```

```{r}
result_time_complete_Q1 
```


### Time in milisec
Create a new column SQ01F01_unit_time by merging the column timeunixsec and timemilisec together by following rule:
the new column SQ01F01_unit_time = timeunixsec * 1000 + timemilisec
```{r}
result_time_complete_Q1$SQ01F01_unit_time <- result_time_complete_Q1$timeunixsec * 1000 + result_time_complete_Q1$timemilisec

# show all the units
result_time_complete_Q1$SQ01F01_unit_time <- format(result_time_complete_Q1$SQ01F01_unit_time, scientific = FALSE)
result_time_complete_Q1
```


### Word Count TA
Filter the data based on PSIVariableIdentifier == 1001 and IsFinalAnswer == 1.
```{r}
filtered_data <- result_time_complete_Q1 %>% filter(PSIVariableIdentifier == 1001)%>%
filter(IsFinalAnswer == 1)
```

```{r}
length(unique(filtered_data$idstud))
```

Create a new column SQ01F01_wordcount_punc_TA that count the number of words in the column information including the punctuations.
```{r}
# Function to split words and punctuations
split_words_and_punctuations <- function(text) {
  # Match words and punctuations separately
  tokens <- strsplit(text, "(?<=[a-zA-Z0-9])(?=[[:punct:]])|(?<=[[:punct:]])(?=[a-zA-Z0-9])|\\s+", perl=TRUE)[[1]]
  # Remove empty tokens
  tokens <- tokens[tokens != ""]
  return(tokens)
}

# Apply the function to the information column
filtered_data$SQ01F01_wordcount_punc_TA <- sapply(filtered_data$information, function(x) length(split_words_and_punctuations(x)))
```
Use (?<=[a-zA-Z0-9])(?=[[:punct:]])|(?<=[[:punct:]])(?=[a-zA-Z0-9]) split words and punctuations separately, and use |\\s+ matches one or more whitespace characters. Then count the total number of tokens.

excluding the punctuations
```{r}
filtered_data <- mutate(filtered_data, SQ01F01_wordcount_no_punc_TA = str_count(gsub("[-[:punct:]]", " ", information), "\\S+"))
filtered_data[c("idstud","information","module_path","SQ01F01_wordcount_punc_TA","SQ01F01_wordcount_no_punc_TA")]
```

Merge the data to the original data
```{r}
# df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01F01_wordcount_punc","SQ01F01_wordcount_no_punc")]
df_filtered_merge = filtered_data[c("idstud","SQ01F01_wordcount_punc_TA","SQ01F01_wordcount_no_punc_TA")]
```

```{r}
# Merge the result back to the original data frame
result_wordcount_Q1_TA <- merge(result_time_complete_Q1, df_filtered_merge, by = "idstud", all.x = TRUE)
result_wordcount_Q1_TA
```


### Raw Response TA
Create a new column SQ01F01_raw_response that is exactly same as the column information of the filtered data.
```{r}
filtered_data$SQ01F01_raw_response_TA <- filtered_data$information

df_filtered_merge = filtered_data[c("idstud","SQ01F01_raw_response_TA")]
```

```{r}
# Merge the result back to the original data frame
result_wordres_Q1_TA <- merge(result_wordcount_Q1_TA, df_filtered_merge, by = "idstud", all.x = TRUE)
result_wordres_Q1_TA
```


### Word Count TB
Filter the data based on PSIVariableIdentifier == 1002 and IsFinalAnswer == 1.
```{r}
filtered_data <- result_time_complete_Q1 %>% filter(PSIVariableIdentifier == 1002)%>%
filter(IsFinalAnswer == 1)
```

```{r}
length(unique(filtered_data$idstud))
```

Create a new column SQ01F01_wordcount_punc that count the number of words in the column information including the punctuations.
```{r}
# Function to split words and punctuations
split_words_and_punctuations <- function(text) {
  # Match words and punctuations separately
  tokens <- strsplit(text, "(?<=[a-zA-Z0-9])(?=[[:punct:]])|(?<=[[:punct:]])(?=[a-zA-Z0-9])|\\s+", perl=TRUE)[[1]]
  # Remove empty tokens
  tokens <- tokens[tokens != ""]
  return(tokens)
}

# Apply the function to the information column
filtered_data$SQ01F01_wordcount_punc_TB <- sapply(filtered_data$information, function(x) length(split_words_and_punctuations(x)))
```
Use (?<=[a-zA-Z0-9])(?=[[:punct:]])|(?<=[[:punct:]])(?=[a-zA-Z0-9]) split words and punctuations separately, and use |\\s+ matches one or more whitespace characters. Then count the total number of tokens.

excluding the punctuations
```{r}
filtered_data <- mutate(filtered_data, SQ01F01_wordcount_no_punc_TB = str_count(gsub("[-[:punct:]]", " ", information), "\\S+"))
filtered_data[c("idstud","information","module_path","SQ01F01_wordcount_punc_TB","SQ01F01_wordcount_no_punc_TB")]
```

Merge the data to the original data
```{r}
# df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01F01_wordcount_punc","SQ01F01_wordcount_no_punc")]
df_filtered_merge = filtered_data[c("idstud","SQ01F01_wordcount_punc_TB","SQ01F01_wordcount_no_punc_TB")]
```

```{r}
# Merge the result back to the original data frame
result_wordcount_Q1_TB <- merge(result_wordres_Q1_TA, df_filtered_merge, by = "idstud", all.x = TRUE)
result_wordcount_Q1_TB
```


### Raw Response TB
Create a new column SQ01F01_raw_response that is exactly same as the column information of the filtered data.
```{r}
filtered_data$SQ01F01_raw_response_TB <- filtered_data$information

df_filtered_merge = filtered_data[c("idstud","SQ01F01_raw_response_TB")]
```

```{r}
# Merge the result back to the original data frame
result_wordres_Q1_TB <- merge(result_wordcount_Q1_TB, df_filtered_merge, by = "idstud", all.x = TRUE)
result_wordres_Q1_TB
```


### All Steps
Create a new dataframe by copy the original one.Create a new column step_str based on the other column.
If PSIVariableIdentifier == 1001, then step_str == "write_explanation_TA"
If PSIVariableIdentifier == 1002, then step_str == "write_explanation_TB"
If eventname == "Nav:GoTo" & new_currentindex == 1, then step_str == "Navigate_to_Page02_Q1"
If eventname == "Nav:GoTo" & new_currentindex != 1, then step_str == if new_currentindex in 1,2,3,5,7,9,10,11,12, then paste0("Navigate_to_Page", new_currentindex+1, "Q",cor_rule(new_currentindex)), else paste0("Navigate_to_Page", new_currentindex+1)
If eventname == "UI:Scrollbar", step_str == extract the value after event and scrollTop in the column information.
If eventname == "Btn:Ruler", step_str == "Ruler" + extract the value after To in the column information.

Create a column new_currentindex by changing the column currentindex based on the rule below:
12 to 0, 13 to 1, 14 to 2, 15 to 3, 16 to 4, 17 to 5, 18 to 6, 19 to 7, 20 to 8, 21 to 9, 22 to 10, 23 to 11, 24 to 12
```{r}
result_copy_Q1=data.frame(result_time_complete_Q1)
result_copy_Q1 <- result_copy_Q1 %>%
  mutate(new_currentindex = case_when(
    currentindex == 12 & BookletPart == 1 ~ 0,
    currentindex == 13 & BookletPart == 1 ~ 1,
    currentindex == 14 ~ 2,
    currentindex == 15 ~ 3,
    currentindex == 16 ~ 4,
    currentindex == 17 ~ 5,
    currentindex == 18 ~ 6,
    currentindex == 19 ~ 7,
    currentindex == 20 ~ 8,
    currentindex == 21 ~ 9,
    currentindex == 22 ~ 10,
    currentindex == 23 ~ 11,
    currentindex == 24 ~ 12,
    currentindex == 25 ~ 13,
    TRUE ~ currentindex  # Keep the original value if it doesn't match any rule
  ))
```

create a cor_rule(new_currentindex) that:
if new_currentindex == 1,2,3,5,7,9,10,11,12, then
if new_currentindex == 1, return 1,
ifelse new_currentindex == 2, return 2,
ifelse new_currentindex == 3, return 3,
ifelse new_currentindex == 5, return 4,
ifelse new_currentindex == 7, return 5,
ifelse new_currentindex == 9, return 6,
ifelse new_currentindex == 10, return 7,
ifelse new_currentindex == 11, return 8,
ifelse new_currentindex == 12, return 9.
```{r}
cor_rule <- function(new_currentindex) {
  result <- case_when(
    new_currentindex == 1 ~ 1,
    new_currentindex == 2 ~ 2,
    new_currentindex == 3 ~ 3,
    new_currentindex == 5 ~ 4,
    new_currentindex == 7 ~ 5,
    new_currentindex == 9 ~ 6,
    new_currentindex == 10 ~ 7,
    new_currentindex == 11 ~ 8,
    new_currentindex == 12 ~ 9,
    TRUE ~ NA_integer_
  )
  return(result)
}
```

if new_currentindex in 1,2,3,5,7,9,10,11,12, then paste0("Navigate_to_Page", new_currentindex+1, "Q",cor_rule(new_currentindex)), else paste0("Navigate_to_Page", new_currentindex+1)

```{r}
result_str_Q1<- result_copy_Q1 %>%
  mutate(
    step_str = case_when(
      PSIVariableIdentifier == 1001 ~ "write_explanation_TA",
      PSIVariableIdentifier == 1002 ~ "write_explanation_TB",
      eventname == "Nav:GoTo" & new_currentindex == 1 ~ "Navigate_to_Page02_Q1",
      eventname == "Nav:GoTo" & new_currentindex != 1 ~ {
        case_when(
          new_currentindex %in% c(1,2,3,5,7,9,10,11,12) ~ paste0("Navigate_to_Page", new_currentindex+1, "_Q", cor_rule(new_currentindex)),
          TRUE ~ paste0("Navigate_to_Page", new_currentindex+1)
        )
      },
      eventname == "UI:Scrollbar" ~ sub('.*"event":"([^"]+).*"scrollTop":(\\d+).*', '\\1 \\2', information),
      eventname == "Btn:Ruler" ~ paste0("Ruler_", gsub('.*To":"(\\w+).*', '\\1', information)),
      TRUE ~ NA_character_  # If none of the conditions are met, set to NA
    )
  )
result_str_Q1[c("RowNumber","step_str","information")]
```
 
If the row of value of the column eventname == "UI:Scrollbar", then the value of the row of the column step_str substitute the white space to _.
```{r}
result_str_use_Q1 <- result_str_Q1 %>%
  mutate(step_str = ifelse(eventname == "UI:Scrollbar", gsub(" ", "_", step_str), step_str))
result_str_use_Q1[c("RowNumber","step_str","information")]
```

```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q1 %>%
  filter(!is.na(step_str)) %>%
  group_by(idstud) %>%
  mutate(SQ01F01_allsteps_list = list(step_str[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01F01_allsteps_seq <- sapply(df_filtered$SQ01F01_allsteps_list, function(x) paste0("(", toString(x), ")"))
```

```{r}
df_filtered <- select(df_filtered, -SQ01F01_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01F01_allsteps_seq")]

result_allsteps_Q1 <- merge(result_wordres_Q1_TB, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_allsteps_Q1=result_allsteps_Q1%>%
  group_by(idstud) %>%
  fill(SQ01F01_allsteps_seq, .direction = "downup")
# result_allsteps_Q1
```

```{r}
result_allsteps_Q1[c("idstud","SQ01F01_allsteps_seq","information")]
```


### Time Taken for All Steps
create a column to record the time taken for each single steps for each student to do this problem.
```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q1 %>%
  group_by(idstud) %>%
  mutate(SQ01F01_allsteps_list = list(SQ01F01_time[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01F01_allsteps_seq_time <- sapply(df_filtered$SQ01F01_allsteps_list, function(x) paste0("(", toString(x), ")"))

# filter the data and select column to merge
df_filtered <- select(df_filtered, -SQ01F01_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01F01_allsteps_seq_time")]

# merge the data
result_alls_time_Q1 <- merge(result_allsteps_Q1, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_alls_time_Q1=result_alls_time_Q1%>%
  group_by(idstud) %>%
  fill(SQ01F01_allsteps_seq_time, .direction = "downup")
result_alls_time_Q1
```


### Timestamp for All Steps in milisecond
create a column to record the time taken for each single steps for each student to do this problem.
```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q1 %>%
  group_by(idstud) %>%
  mutate(SQ01F01_allsteps_list = list(SQ01F01_unit_time[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01F01_allsteps_timestamp <- sapply(df_filtered$SQ01F01_allsteps_list, function(x) paste0("(", toString(x), ")"))

# filter the data and select column to merge
df_filtered <- select(df_filtered, -SQ01F01_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01F01_allsteps_timestamp")]

# merge the data
result_all_unit_time_Q1 <- merge(result_alls_time_Q1, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_all_unit_time_Q1=result_all_unit_time_Q1%>%
  group_by(idstud) %>%
  fill(SQ01F01_allsteps_timestamp, .direction = "downup")
result_all_unit_time_Q1
```

### Total time taken
Create a new column SQ01F01_total_time. Group by idstud, SQ01F01_total_time = sum the value of the column SQ01F01_time for each idstud.

```{r}
# Group by idstud and calculate the sum of SQ01F01_time
result <- result_time_complete_Q1 %>%
  group_by(idstud) %>%
  summarise(SQ01F01_total_time = sum(SQ01F01_time, na.rm = TRUE))

# Merge the result back to the original data frame
result_total_time_Q1 <- merge(result_all_unit_time_Q1, result, by = "idstud", all.x = TRUE)
```

```{r}
# result_total_time_Q1[c("idstud","SQ01F01_time","SQ01F01_total_time")]
result_total_time_Q1
```


### Save the data
Filter the columns of data frame, only keep the new created columns and idstud, idstud_ori, ImportLogID, BlockName, BookletPart, BlockInPart.
```{r}
resultF01 <- result_total_time_Q1 %>%
  select(idstud, idstud_ori, ImportLogID, BlockName, BookletPart,BlockInPart, module_path, SQ01F01_wordcount_punc_TA, SQ01F01_wordcount_no_punc_TA, SQ01F01_raw_response_TA, SQ01F01_wordcount_punc_TB, SQ01F01_wordcount_no_punc_TB, SQ01F01_raw_response_TB, SQ01F01_allsteps_seq, SQ01F01_allsteps_seq_time, SQ01F01_allsteps_timestamp, SQ01F01_total_time)
# resultF01
```


```{r}
resultF01_save = resultF01[!duplicated(resultF01), ]

condition1 <- resultF01_save$BlockName == 'SQ01' & 
              resultF01_save$BookletPart == 1 & 
              resultF01_save$BlockInPart == 2 

condition2 <- resultF01_save$BlockName == 'SQ01' & 
              resultF01_save$BookletPart == 2 & 
              resultF01_save$BlockInPart == 1

combined_condition <- condition1 | condition2

resultF01_save=resultF01_save%>%ungroup()%>% 
  filter(combined_condition)
resultF01_save
```

Save the dataframe
```{r}
# write_sav(resultF01_save, "../../../new_data/grade5/Farm/SQ01F01_PSI.sav")

# write.csv(resultF01_save, file = 'csv_SQ01F01_PSI.csv', row.names = FALSE)
```






