---
title: "F08"
author: "zifengxu"
date: "2024-04-17"
output: html_document
---

# Read the Data (PSI)
```{r}
library(haven)
library(dplyr)
library(stringr)
library(tidyr)
```

```{r}
df <- read_sav("../../../data/grade5/ASLNORZ7_processdata_PSIstudents.sav")

# Explore the data
head(df)
```

Remove the rows of auto check for this question
```{r}
df_filter = df%>%
  filter(df$PSIVariableIdentifier != 1112 |is.na(df$PSIVariableIdentifier))
```

Fiter the rows and exactly one row above the row that satisfied one of the conditions below: 

Condition 1: BlockName=SQ01, BookletPart=1, BlockInPart=2, currentindex=23.

Condition 2: BlockName=SQ01, BookletPart=2, BlockInPart=1, currentindex=11.

```{r}
condition1 <- df_filter$BlockName == 'SQ01' & 
              df_filter$BookletPart == 1 & 
              df_filter$BlockInPart == 2 & 
              df_filter$currentindex == 23 

# Define the second condition
condition2 <- df_filter$BlockName == 'SQ01' & 
              df_filter$BookletPart == 2 & 
              df_filter$BlockInPart == 1 & 
              df_filter$currentindex == 11 

# Combine the two conditions using OR (|) as you want rows satisfying either condition
combined_condition <- condition1 | condition2

# Use filter to get the rows that satisfy the conditions and one row below
result8 <- df_filter %>% 
  filter(combined_condition | lag(combined_condition, default = FALSE))
```

```{r}
result8
```

```{r}
result8[c("idstud","eventname","information","PSIVariableIdentifier","IsFinalAnswer")]
```


### Module Path
Create a new column module_path based on (BlockName, BookletPart, BlockInPart, currentindex)
```{r}
result_use_Q8= transform(result8,
                       module_path = paste0("(", BlockName, ", ", BookletPart, ", ", BlockInPart, ")"))
# result_use_Q8
```


### Time
```{r}
result_filter_Q8=result_use_Q8 %>%filter(eventname %in% c("Nav:GoTo", "Response", "UI:Scrollbar", "Btn:Ruler"))
result_filter_Q8
```

```{r}
result_time_complete_Q8=result_filter_Q8 %>%
  mutate(SQ01F08_time = ifelse(currentindex == 23 & eventname == 'Nav:GoTo' | currentindex == 11 & eventname == 'Nav:GoTo', 0, ifelse(timemilisec - lag(timemilisec) > 0,
                                      (timeunixsec - lag(timeunixsec) + (timemilisec - lag(timemilisec))/1000),
                                      (timeunixsec - lag(timeunixsec) - 1 + (timemilisec + 1000 - lag(timemilisec))/1000))))

# head(result_time_complete_Q8[c("idstud","module_path","SQ01F08_time")])
# result_time_complete_Q8
```

```{r}
result_time_complete_Q8 = result_time_complete_Q8%>%
  mutate(RowNumber = row_number())
```

```{r}
result_time_complete_Q8 
```

```{r}
result_time_complete_Q8$SQ01F08_unit_time <- result_time_complete_Q8$timeunixsec * 1000 + result_time_complete_Q8$timemilisec

# show all the units
result_time_complete_Q8$SQ01F08_unit_time <- format(result_time_complete_Q8$SQ01F08_unit_time, scientific = FALSE)
result_time_complete_Q8
```


### Blue Circle Sequence A
Create a new column SQ01F08_A_slider_seq to record the sequence that student slide the blue circle of Part A based on the following conditions: Filter the data by using PSIVariableIdentifier == 1110, which means the students move the blue circle. Then group by the student id, the new column SQ01F08_A_slider_seq should created based on the sequence of the value of the column information in the order of if the row number is higher, their sequence should be higher.

```{r}
# Filter data and arrange by idstud and information
df_filtered <- result_time_complete_Q8 %>%
  filter(PSIVariableIdentifier == 1110)%>%
  group_by(idstud) %>%
  mutate(SQ01F08_sequence_list = list(information[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01F08_A_slider_seq <- sapply(df_filtered$SQ01F08_sequence_list, function(x) paste0("(", toString(x), ")"))

df_filtered
```

Create a new dataframe by merge the two dataframes by using the column idstud, and RowNumber. 
```{r}
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01F08_A_slider_seq")]
```

```{r}
result_A_seq_Q8 <- merge(result_time_complete_Q8, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)
# head(result_A_seq_Q8)
```

Then fill in all the missing value of the column SQ01F08_A_slider_seq according to the student name.
```{r}
result_A_seq_Q8=result_A_seq_Q8%>%
  group_by(idstud) %>%
  fill(SQ01F08_A_slider_seq, .direction = "downup")
result_A_seq_Q8
```

Check the missing value, there exists missing value because some students do not do this step.
```{r}
# result_A_seq_Q8[is.na(result_A_seq_Q8$SQ01F08_sequence), ]
```


### Final Slider Position A
Create a new column SQ01F08_A_final_pos. Filter the data by using PSIVariableIdentifier == 1110 and IsfinalAnswer ==1, then groupby idstud, SQ01F08_A_final_pos = information.
```{r}
# Filter the data based on PSIVariableIdentifier == 1110
filtered_data <- result_time_complete_Q8 %>% filter(PSIVariableIdentifier == 1110)%>%
filter(IsFinalAnswer == 1)

# Group by idstud and create the new column SQ01F08_A_final_pos
result <- filtered_data %>%
  group_by(idstud) %>%
  summarise(SQ01F08_A_final_pos = information)

# Merge the result back to the original data frame
result_A_Final_pos_Q8 <- merge(result_A_seq_Q8, result, by = "idstud", all.x = TRUE)

result_A_Final_pos_Q8
```

### Time in between the Sequence A
Then record the time taken for each time they move the blue circle. Filter by PSIVariableIdentifier == 1110 which means the students move the blue circle. Then group by the student id, the new column SQ01F08_A_slider_seq_time should created based on the sequence of the value of the column SQ01F08_time in the order of if the row number is higher, their sequence should be higher.
```{r}
# Filter data and arrange by idstud and SQ01F08_time
df_filtered <- result_time_complete_Q8 %>%
  filter(PSIVariableIdentifier == 1110)%>%
  group_by(idstud) %>%
  mutate(SQ01F08_seq_time_list = list(SQ01F08_time[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01F08_A_slider_seq_time <- sapply(df_filtered$SQ01F08_seq_time_list, function(x) paste0("(", toString(x), ")"))

df_filtered
```


```{r}
# create a new dataframe to merge
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01F08_A_slider_seq_time")]

# merge the data
result_A_seqtime_Q8 <- merge(result_A_Final_pos_Q8, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

# fill the missing value of the column SQ01F08_sequence based on the student name
result_A_seqtime_Q8=result_A_seqtime_Q8%>%
  group_by(idstud) %>%
  fill(SQ01F08_A_slider_seq_time, .direction = "downup")
result_A_seqtime_Q8
```


### Blue Circle Sequence B
Create a new column SQ01F08_B_slider_seq to store the sequence that student slide the blue circle of Part B based on the following conditions:
Filter the data by using PSIVariableIdentifier == 1111, which means the students move the blue circle. Then group by the student id, the new column SQ01F08_B_slider_seq should created based on the sequence of the value of the column information in the order of if the row number is higher, their sequence should be higher.

```{r}
# Filter data and arrange by idstud and information
df_filtered <- result_time_complete_Q8 %>%
  filter(PSIVariableIdentifier == 1111)%>%
  group_by(idstud) %>%
  mutate(SQ01F08_sequence_list = list(information[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01F08_B_slider_seq <- sapply(df_filtered$SQ01F08_sequence_list, function(x) paste0("(", toString(x), ")"))

df_filtered
```

Create a new dataframe by merge the two dataframes by using the column idstud, and RowNumber. 
```{r}
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01F08_B_slider_seq")]
```

```{r}
result_B_seq_Q8 <- merge(result_A_seqtime_Q8, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)
# head(result_B_seq_Q8)
```

Then fill in all the missing value of the column SQ01F08_B_slider_seq according to the student name.
```{r}
result_B_seq_Q8=result_B_seq_Q8%>%
  group_by(idstud) %>%
  fill(SQ01F08_B_slider_seq, .direction = "downup")
result_B_seq_Q8
```

Check the missing value, there exists missing value because some students do not do this step.
```{r}
# result_B_seq_Q8[is.na(result_B_seq_Q8$SQ01F08_sequence), ]
```


### Final Slider Position B
Create a new column SQ01F08_A_final_pos. Filter the data by using PSIVariableIdentifier == 1111 and IsfinalAnswer ==1, then groupby idstud, SQ01F08_A_final_pos = information.
```{r}
# Filter the data based on PSIVariableIdentifier == 1110
filtered_data <- result_time_complete_Q8 %>% filter(PSIVariableIdentifier == 1111)%>%
filter(IsFinalAnswer == 1)

# Group by idstud and create the new column SQ01F08_A_final_pos
result <- filtered_data %>%
  group_by(idstud) %>%
  summarise(SQ01F08_B_final_pos = information)

# Merge the result back to the original data frame
result_B_Final_pos_Q8 <- merge(result_B_seq_Q8, result, by = "idstud", all.x = TRUE)

result_B_Final_pos_Q8
```


### Time in between the Sequence B
Then record the time taken for each time they move the blue circle. Filter by PSIVariableIdentifier == 1111 which means the students move the blue circle. Then group by the student id, the new column SQ01F08_B_slider_seq_time should created based on the sequence of the value of the column SQ01F08_time in the order of if the row number is higher, their sequence should be higher.
```{r}
# Filter data and arrange by idstud and SQ01F08_time
df_filtered <- result_time_complete_Q8 %>%
  filter(PSIVariableIdentifier == 1111)%>%
  group_by(idstud) %>%
  mutate(SQ01F08_seq_time_list = list(SQ01F08_time[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01F08_B_slider_seq_time <- sapply(df_filtered$SQ01F08_seq_time_list, function(x) paste0("(", toString(x), ")"))

df_filtered
```


```{r}
# create a new dataframe to merge
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01F08_B_slider_seq_time")]

# merge the data
result_B_seqtime_Q8 <- merge(result_B_Final_pos_Q8, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

# fill the missing value of the column SQ01F08_sequence based on the student name
result_B_seqtime_Q8=result_B_seqtime_Q8%>%
  group_by(idstud) %>%
  fill(SQ01F08_B_slider_seq_time, .direction = "downup")
result_B_seqtime_Q8
```


### All Steps
Create a new dataframe by copy the original one.Create a new column step_str based on the other column.
If PSIVariableIdentifier == 1110, then step_str == paste0("sliderA_",information)
If PSIVariableIdentifier == 1111, then step_str == paste0("sliderB_",information)
If eventname == "Nav:GoTo" & new_currentindex == 11, then step_str == "Navigate_to_Page12_Q8"
If eventname == "Nav:GoTo" & new_currentindex != 11, then step_str == if new_currentindex in 1,2,3,5,7,9,10,11,12, then paste0("Navigate_to_Page", new_currentindex+1, "Q",cor_rule(new_currentindex)), else paste0("Navigate_to_Page", new_currentindex+1)
If eventname == "UI:Scrollbar", step_str == extract the value after event and scrollTop in the column information.
If eventname == "Btn:Ruler", step_str == "Ruler" + extract the value after To in the column information.

Create a column new_currentindex by changing the column currentindex based on the rule below:
15 to 0, 16 to 1, 17 to 2, 18 to 3, 19 to 4, 20 to 56, 21 to 6, 22 to 7, 23 to 8, 24 to 9, 25 to 10, 26 to 11, 27 to 12, 28 to 13
```{r}
result_copy_Q8=data.frame(result_time_complete_Q8)
result_copy_Q8 <- result_copy_Q8 %>%
  mutate(new_currentindex = case_when(
    currentindex == 12 & BookletPart == 1 ~ 0,
    currentindex == 13 & BookletPart == 1 ~ 1,
    currentindex == 14 ~ 2,
    currentindex == 15 ~ 3,
    currentindex == 16 ~ 4,
    currentindex == 17 ~ 5,
    currentindex == 18 ~ 6,
    currentindex == 19 ~ 7,
    currentindex == 20 ~ 8,
    currentindex == 21 ~ 9,
    currentindex == 22 ~ 10,
    currentindex == 23 ~ 11,
    currentindex == 24 ~ 12,
    currentindex == 25 ~ 13,
    TRUE ~ currentindex  # Keep the original value if it doesn't match any rule
  ))
```

create a cor_rule(new_currentindex) that:
if new_currentindex == 1,2,3,5,7,9,10,11,12, then
if new_currentindex == 1, return 1,
ifelse new_currentindex == 2, return 2,
ifelse new_currentindex == 3, return 3,
ifelse new_currentindex == 5, return 4,
ifelse new_currentindex == 7, return 5,
ifelse new_currentindex == 9, return 6,
ifelse new_currentindex == 10, return 7,
ifelse new_currentindex == 11, return 8,
ifelse new_currentindex == 12, return 9.
```{r}
cor_rule <- function(new_currentindex) {
  result <- case_when(
    new_currentindex == 1 ~ 1,
    new_currentindex == 2 ~ 2,
    new_currentindex == 3 ~ 3,
    new_currentindex == 5 ~ 4,
    new_currentindex == 7 ~ 5,
    new_currentindex == 9 ~ 6,
    new_currentindex == 10 ~ 7,
    new_currentindex == 11 ~ 8,
    new_currentindex == 12 ~ 9,
    TRUE ~ NA_integer_
  )
  return(result)
}
```

```{r}
result_str_Q8<- result_copy_Q8 %>%
  mutate(
    step_str = case_when(
      PSIVariableIdentifier == 1110 ~ paste0("sliderA_",information),
      PSIVariableIdentifier == 1111 ~ paste0("sliderB_",information),
      eventname == "Nav:GoTo" & new_currentindex == 11 ~ "Navigate_to_Page12_Q8",
      eventname == "Nav:GoTo" & new_currentindex != 11 ~ {
        case_when(
          new_currentindex %in% c(1,2,3,5,7,9,10,11,12) ~ paste0("Navigate_to_Page", new_currentindex+1, "_Q", cor_rule(new_currentindex)),
          TRUE ~ paste0("Navigate_to_Page", new_currentindex+1)
        )
      },
      eventname == "UI:Scrollbar" ~ sub('.*"event":"([^"]+).*"scrollTop":(\\d+).*', '\\1 \\2', information),
      eventname == "Btn:Ruler" ~ paste0("Ruler_", gsub('.*To":"(\\w+).*', '\\1', information)),
      TRUE ~ NA_character_  # If none of the conditions are met, set to NA
    )
  )
result_str_Q8[c("RowNumber","step_str","information")]
```
 
If the row of value of the column eventname == "UI:Scrollbar", then the value of the row of the column step_str substitute the white space to _.
```{r}
result_str_use_Q8 <- result_str_Q8 %>%
  mutate(step_str = ifelse(eventname == "UI:Scrollbar", gsub(" ", "_", step_str), step_str))
result_str_use_Q8[c("RowNumber","step_str","information")]
```

```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q8 %>%
  filter(!is.na(step_str)) %>%
  group_by(idstud) %>%
  mutate(SQ01F08_allsteps_list = list(step_str[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01F08_allsteps_seq <- sapply(df_filtered$SQ01F08_allsteps_list, function(x) paste0("(", toString(x), ")"))
```

```{r}
df_filtered <- select(df_filtered, -SQ01F08_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01F08_allsteps_seq")]

result_allsteps_Q8 <- merge(result_B_seqtime_Q8, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_allsteps_Q8=result_allsteps_Q8%>%
  group_by(idstud) %>%
  fill(SQ01F08_allsteps_seq, .direction = "downup")

result_allsteps_Q8
```

```{r}
result_allsteps_Q8[c("idstud","SQ01F08_allsteps_seq","information")]
```


### Time Taken for All Steps
create a column to record the time taken for each single steps for each student to do this problem.
```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q8 %>%
  group_by(idstud) %>%
  mutate(SQ01F08_allsteps_list = list(SQ01F08_time[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01F08_allsteps_seq_time <- sapply(df_filtered$SQ01F08_allsteps_list, function(x) paste0("(", toString(x), ")"))

# filter the data and select column to merge
df_filtered <- select(df_filtered, -SQ01F08_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01F08_allsteps_seq_time")]

# merge the data
result_alls_time_Q8 <- merge(result_allsteps_Q8, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_alls_time_Q8=result_alls_time_Q8%>%
  group_by(idstud) %>%
  fill(SQ01F08_allsteps_seq_time, .direction = "downup")
result_alls_time_Q8
```


### Timestamp for All Steps in milisecond
create a column to record the time taken for each single steps for each student to do this problem.
```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q8 %>%
  group_by(idstud) %>%
  mutate(SQ01F08_allsteps_list = list(SQ01F08_unit_time[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01F08_allsteps_timestamp <- sapply(df_filtered$SQ01F08_allsteps_list, function(x) paste0("(", toString(x), ")"))

# filter the data and select column to merge
df_filtered <- select(df_filtered, -SQ01F08_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01F08_allsteps_timestamp")]

# merge the data
result_all_unit_time_Q8 <- merge(result_alls_time_Q8, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_all_unit_time_Q8=result_all_unit_time_Q8%>%
  group_by(idstud) %>%
  fill(SQ01F08_allsteps_timestamp, .direction = "downup")
result_all_unit_time_Q8
```

### Total time taken
Create a new column SQ01F08_total_time. Group by idstud, SQ01F08_total_time = sum the value of the column SQ01F08_time for each idstud.

```{r}
# Group by idstud and calculate the sum of SQ01F08_time
result <- result_time_complete_Q8 %>%
  group_by(idstud) %>%
  summarise(SQ01F08_total_time = sum(SQ01F08_time, na.rm = TRUE))

# Merge the result back to the original data frame
result_total_time_Q8 <- merge(result_all_unit_time_Q8, result, by = "idstud", all.x = TRUE)
```

```{r}
result_total_time_Q8
# result_total_time_Q8[c("idstud","SQ01F08_time","SQ01F08_total_time")]
```


### Save the data
Filter the columns of data frame, only keep the new created columns and idstud, idstud_ori, ImportLogID, BlockName, BookletPart, BlockInPart.
```{r}
resultF08 <- result_total_time_Q8 %>%
  select(idstud, idstud_ori, ImportLogID, BlockName, BookletPart,BlockInPart, module_path, SQ01F08_A_slider_seq, SQ01F08_A_final_pos, SQ01F08_A_slider_seq_time, SQ01F08_B_slider_seq, SQ01F08_B_final_pos, SQ01F08_B_slider_seq_time, SQ01F08_allsteps_seq, SQ01F08_allsteps_seq_time, SQ01F08_allsteps_timestamp, SQ01F08_total_time)
# resultF08
```


```{r}
resultF08_save = resultF08[!duplicated(resultF08), ]

condition1 <- resultF08_save$BlockName == 'SQ01' & 
              resultF08_save$BookletPart == 1 & 
              resultF08_save$BlockInPart == 2 

condition2 <- resultF08_save$BlockName == 'SQ01' & 
              resultF08_save$BookletPart == 2 & 
              resultF08_save$BlockInPart == 1

combined_condition <- condition1 | condition2

resultF08_save=resultF08_save%>%ungroup()%>% 
  filter(combined_condition)
resultF08_save
```

Save the dataframe
```{r}
# write_sav(resultF08_save, "../../../new_data/grade5/Farm/SQ01F08_PSI.sav")

# write.csv(resultF08_save, file = 'csv_SQ01F08_PSI.csv', row.names = FALSE)
```




