---
title: "SQ02all"
author: "zifengxu"
date: "2024-03-28"
output: html_document
---

# Read the Data (PSI)
```{r}
library(haven)
library(dplyr)
library(stringr)
library(tidyr)
library(jsonlite)
```

```{r}
df <- read_sav("../../data/grade9/BSLNORZ7_PSI.sav")

# Explore the data
head(df)
```

```{r}
df_filter = df%>%
  filter(eventname != "Btn:Next")
```

Filter the rows and exactly one row above the row that satisfied one of the conditions below: 

Condition 1: BlockName=SQ02, BookletPart=1, BlockInPart=1.

Condition 2: BlockName=SQ02, BookletPart=2, BlockInPart=2.

```{r}
condition1 <- df_filter$BlockName == 'SQ02' & 
              df_filter$BookletPart == 1 & 
              df_filter$BlockInPart == 1  
 

condition2 <- df_filter$BlockName == 'SQ02' & 
              df_filter$BookletPart == 2 & 
              df_filter$BlockInPart == 2  


# Combine the two conditions using OR (|) as you want rows satisfying either condition
combined_condition <- condition1 | condition2

# Use filter to get the rows that satisfy the conditions and one row below
result_all <- df_filter %>% 
  filter(combined_condition | lag(combined_condition, default = FALSE))

result_all
```

Filter all the rows that 
(1)the value of the column currentindex or the value of the column idstud is different from the previous row.
(2)the first row
```{r}
result_all_use = result_all%>%
  filter(row_number() == 1 | 
         currentindex != lag(currentindex) | 
         idstud != lag(idstud) | 
         eventname == "Nav:GoTo")
result_all_use
```

### Time
Create a new column SQ02_time based on the following rule: the new column SQ02_time =
ifelse(lead(timemilisec) - timemilisec > 0, (lead(timeunixsec)-timeunixsec + (lead(timemilisec)-timemilisec)/1000),(lead(timeunixsec)-timeunixsec - 1 + (lead(timemilisec)+1000 -timemilisec)/1000))
```{r}
result_time_complete_all=result_all_use %>%
  arrange(idstud, timeunixsec, timemilisec) %>%
  group_by(idstud) %>%
  mutate(
    SQ02_time =ifelse(lead(timemilisec) - timemilisec > 0, (lead(timeunixsec)-timeunixsec + (lead(timemilisec)-timemilisec)/1000),(lead(timeunixsec)-timeunixsec - 1 + (lead(timemilisec)+1000 -timemilisec)/1000))
  )
result_time_complete_all
```

```{r}
result_time_complete_all[, c('idstud', 'SQ02_time', 'timeunixsec', 'timemilisec')]
```

```{r}
result_time_complete_all = result_time_complete_all%>%
  mutate(RowNumber = row_number())
```

```{r}
result_time_complete_all$SQ02_all_unit_time <- result_time_complete_all$timeunixsec * 1000 + result_time_complete_all$timemilisec

# show all the units
result_time_complete_all$SQ02_all_unit_time <- format(result_time_complete_all$SQ02_all_unit_time, scientific = FALSE)
```

```{r}
result_time_complete_all[, c('idstud', 'SQ02_time','currentindex' ,'SQ02_all_unit_time', 'timeunixsec', 'timemilisec')]
# result_time_complete_all
```

### Filter the using data
Filter the data based on the condition below:
currentindex == 0,1,2,3,4,5,6,7,8,9,10,11,12,13,15,16,17,18,19,20,21,22,23,24,25,26,27,28 & eventname == "Nav:GoTo" or "	UI:IsLoaded" & BlockName = "SQ02"
```{r}
result_time_complete_all=result_time_complete_all%>%
  filter(
    currentindex %in% c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25, 26, 27, 28) &
    (eventname == "Nav:GoTo" | eventname == "UI:IsLoaded") &
    BlockName == "SQ02"
  )
result_time_complete_all
```


### All sequence
```{r}
result_copy_all=data.frame(result_time_complete_all)
result_copy_all <- result_copy_all %>%
  mutate(new_currentindex = case_when(
    currentindex == 15 ~ 1,
    currentindex == 16 ~ 2,
    currentindex == 17 ~ 3,
    currentindex == 18 ~ 4,
    currentindex == 19 ~ 5,
    currentindex == 20 ~ 6,
    currentindex == 21 ~ 7,
    currentindex == 22 ~ 8,
    currentindex == 23 ~ 9,
    currentindex == 24 ~ 10,
    currentindex == 25 ~ 11,
    currentindex == 26 ~ 12,
    currentindex == 27 ~ 13,
    currentindex == 28 ~ 14,
    TRUE ~ currentindex+1  
  ))
```


```{r}
df_filtered <- result_copy_all %>%
  group_by(idstud) %>%
  mutate(SQ02_Navigate_list = list(new_currentindex[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ02_Navigate_seq <- sapply(df_filtered$SQ02_Navigate_list, function(x) paste0("(", toString(x), ")"))
```

```{r}
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ02_Navigate_seq")]

result_Navi_all <- merge(result_time_complete_all, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_Navi_all=result_Navi_all%>%
  group_by(idstud) %>%
  fill(SQ02_Navigate_seq, .direction = "downup")

result_Navi_all
```


### Time Taken for All Steps
create a column to record the time taken for each single steps for each student to do this problem.
```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_copy_all %>%
  group_by(idstud) %>%
  mutate(SQ02_Navigate_list = list(SQ02_time[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ02_Navigate_seq_time <- sapply(df_filtered$SQ02_Navigate_list, function(x) paste0("(", toString(x), ")"))

# filter the data and select column to merge
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ02_Navigate_seq_time")]

# merge the data
result_alls_time_all <- merge(result_Navi_all, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_alls_time_all=result_alls_time_all%>%
  group_by(idstud) %>%
  fill(SQ02_Navigate_seq_time, .direction = "downup")
result_alls_time_all
```


### Timestamp for All Steps in milisecond
```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_copy_all %>%
  group_by(idstud) %>%
  mutate(SQ02_Navigate_list = list(SQ02_all_unit_time[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ02_Navigate_timestamp <- sapply(df_filtered$SQ02_Navigate_list, function(x) paste0("(", toString(x), ")"))

# filter the data and select column to merge
df_filtered <- select(df_filtered, -SQ02_Navigate_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ02_Navigate_timestamp")]

# merge the data
result_all_unit_time_all <- merge(result_alls_time_all, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_all_unit_time_all=result_all_unit_time_all%>%
  group_by(idstud) %>%
  fill(SQ02_Navigate_timestamp, .direction = "downup")
result_all_unit_time_all
```

### Total time taken
Create a new column SQ02_unit_total_time. Group by idstud, SQ02_unit_total_time = sum the value of the column SQ02_time for each idstud.
```{r}
# Group by idstud and calculate the sum of SQ02_time
result <- result_copy_all %>%
  group_by(idstud) %>%
  summarise(SQ02_unit_total_time = sum(SQ02_time, na.rm = TRUE))

# Merge the result back to the original data frame
result_total_time_all <- merge(result_all_unit_time_all, result, by = "idstud", all.x = TRUE)
```

```{r}
result_total_time_all
# result_total_time_all[c("idstud","SQ02_time","SQ02_unit_total_time")]
```


### Save the data
Filter the columns of data frame, only keep the new created columns and idstud, idstud_ori, ImportLogID, BlockName, BookletPart, BlockInPart.
```{r}
resultAll <- result_total_time_all %>%
  select(idstud, idstud_ori, ImportLogID, BlockName, BookletPart,BlockInPart, SQ02_Navigate_seq, SQ02_Navigate_seq_time, SQ02_Navigate_timestamp, SQ02_unit_total_time)
# resultAll
```


```{r}
resultAll_save = resultAll[!duplicated(resultAll), ]

condition1 <- resultAll_save$BlockName == 'SQ02' & 
              resultAll_save$BookletPart == 1 & 
              resultAll_save$BlockInPart == 1 

condition2 <- resultAll_save$BlockName == 'SQ02' & 
              resultAll_save$BookletPart == 2 & 
              resultAll_save$BlockInPart == 2

combined_condition <- condition1 | condition2

resultAll_save=resultAll_save%>%ungroup()%>% 
  filter(combined_condition)
resultAll_save
```

Save the dataframe
```{r}
# write_sav(resultAll_save, "../../../new_data/grade9/Pepper_Plants/SQ02Pall_PSI.sav")

# write.csv(resultAll_save, file = 'csv_SQ12all_PSI.csv', row.names = FALSE)
```


