---
title: "SQ01_S03"
author: "zifengxu"
date: "2023-12-11"
output: html_document
---

# Read the Data (PSI)
```{r}
library(haven)
library(dplyr)
library(stringr)
library(tidyr)
library(jsonlite)
```

```{r}
df <- read_sav("../data/grade9/BSLNORZ7_PSI.sav")

# Explore the data
head(df)
```

Remove the rows of auto check for this question
```{r}
df_filter = df%>%
  filter(df$PSIVariableIdentifier != 5013 |is.na(df$PSIVariableIdentifier))
```

Fiter the rows and exactly one row above the row that satisfied one of the conditions below: 

Condition 1: BlockName=SQ01, BookletPart=1, BlockInPart=2, currentindex=20.

Condition 2: BlockName=SQ01, BookletPart=2, BlockInPart=1, currentindex=6.

```{r}
condition1 <- df_filter$BlockName == 'SQ01' & 
              df_filter$BookletPart == 1 & 
              df_filter$BlockInPart == 2 & 
              df_filter$currentindex == 17 

# Define the second condition
condition2 <- df_filter$BlockName == 'SQ01' & 
              df_filter$BookletPart == 2 & 
              df_filter$BlockInPart == 1 & 
              df_filter$currentindex == 3 

# Combine the two conditions using OR (|) as you want rows satisfying either condition
combined_condition <- condition1 | condition2

# Use filter to get the rows that satisfy the conditions and one row below
result3 <- df_filter %>% 
  filter(combined_condition | lag(combined_condition, default = FALSE))

result3
```


### Module Path
Create a new column module_path based on (BlockName, BookletPart, BlockInPart, currentindex)
```{r}
result_use_Q3= transform(result3,
                       module_path = paste0("(", BlockName, ", ", BookletPart, ", ", BlockInPart, ")"))
head(result_use_Q3)
```

```{r}
unique(result_use_Q3$eventname)
```


### Time
```{r}
result_filter_Q3=result_use_Q3 %>%filter(eventname %in% c("Nav:GoTo", "Response", "UI:Scrollbar", "Btn:Calc"))
result_filter_Q3
```


```{r}
result_time_complete_Q3=result_filter_Q3 %>%
  mutate(SQ01S03_time = ifelse(currentindex == 17 & eventname == 'Nav:GoTo' | currentindex == 3 & eventname == 'Nav:GoTo', 0, ifelse(timemilisec - lag(timemilisec) > 0,
                                      (timeunixsec - lag(timeunixsec) + (timemilisec - lag(timemilisec))/1000),
                                      (timeunixsec - lag(timeunixsec) - 1 + (timemilisec + 1000 - lag(timemilisec))/1000))))
# head(result_time_complete_Q3[c("idstud","module_path","SQ01S03_time")])

# result_time_complete_Q3
```

```{r}
result_time_complete_Q3 = result_time_complete_Q3%>%
  mutate(RowNumber = row_number())
```

```{r}
result_time_complete_Q3$SQ01S03_unit_time <- result_time_complete_Q3$timeunixsec * 1000 + result_time_complete_Q3$timemilisec

# show all the units
result_time_complete_Q3$SQ01S03_unit_time <- format(result_time_complete_Q3$SQ01S03_unit_time, scientific = FALSE)
```

```{r}
result_time_complete_Q3
```


### Count how many time undo
```{r}
df_check =result_time_complete_Q3[result_time_complete_Q3$PSIVariableIdentifier== 5010,]
df_check[c("idstud","eventname","information")]
```

Create a new column SQ01S03_undo. Filter the data by using PSIVariableIdentifier == 5010, then groupby idstud, for each different idstud, count how many time the column information has the value of 0.

```{r}
# Create a new column SQ01S03_undo
df_filtered <- result_time_complete_Q3%>% mutate(SQ01S03_undo = NA)

# Filter the data based on PSIVariableIdentifier == 5010
filtered_data <- df_filtered %>% filter(PSIVariableIdentifier == 5010)

# Group by idstud and count occurrences of 0 in the information column
result <- filtered_data %>%
  group_by(idstud) %>%
  summarise(SQ01S03_undo = sum(information == 0, na.rm = TRUE))

# Merge the result back to the original data frame
result_undo <- merge(result_time_complete_Q3, result, by = "idstud", all.x = TRUE)

result_undo
```


### Check if the student drop the coin.
```{r}
df_check=result_undo%>%
  filter(PSIVariableIdentifier == 5010)
df_check[c("idstud","eventname","information")]
```

Create a new column SQ01S03_drop_coin. Filter the data by using PSIVariableIdentifier == 5010, then groupby idstud, for each different idstud, if at least one time the information column has the value = 1, then return 1, otherwise return 0.
```{r}
# Filter the data based on PSIVariableIdentifier == 5010
filtered_data <- result_undo %>% filter(PSIVariableIdentifier == 5010)

# Group by idstud and check if at least one occurrence of information is 1
result <- filtered_data %>%
  group_by(idstud) %>%
  summarise(SQ01S03_drop_coin = any(information == 1, na.rm = TRUE))

# Merge the result back to the original data frame
result_dropcoin_Q3 <- merge(result_undo, result, by = "idstud", all.x = TRUE)

# If there are NAs in the new column, replace them with 0
result_dropcoin_Q3$SQ01S03_drop_coin[is.na(result_dropcoin_Q3$SQ01S03_drop_coin)] <- 0
```

```{r}
result_dropcoin_Q3
# result_dropcoin_Q3[c("idstud","RowNumber","information","SQ01S03_drop_coin")]
```


### Final Answer
```{r}
df_check =result_time_complete_Q3[result_time_complete_Q3$PSIVariableIdentifier== 5011,]
df_check[c("idstud","eventname","IsFinalAnswer","information")]
```

Create a new column SQ01S03_final_answer. Filter the data by using PSIVariableIdentifier == 5011 and IsfinalAnswer ==1, then groupby idstud, SQ01S03_final_answer = information.
```{r}
# Filter the data based on PSIVariableIdentifier == 5011
filtered_data <- result_time_complete_Q3 %>% filter(PSIVariableIdentifier == 5011)%>%
filter(IsFinalAnswer == 1)

# Group by idstud and create the new column SQ01S03_final_answer
result <- filtered_data %>%
  group_by(idstud) %>%
  summarise(SQ01S03_final_answer = information)

# Merge the result back to the original data frame
result_Final_Answer_Q3 <- merge(result_dropcoin_Q3, result, by = "idstud", all.x = TRUE)
```

```{r}
result_Final_Answer_Q3
# result_Final_Answer_Q3[c("idstud","eventname","IsFinalAnswer","PSIVariableIdentifier","information","SQ01S03_final_answer")]
```


### Check how many times come back to the previous question.
Create a new column SQ01S03_goback. Filter the data according to BookletPart=1, BlockInPart=2, currentindex=16, eventname="Nav:GoTo" or BookletPart=2, BlockInPart=1, currentindex=2, eventname="Nav:GoTo". Count how the occurance for each different value of the column idstud.

```{r}
# Filter the data based on the specified conditions
filtered_data <- result_time_complete_Q3 %>%
  filter(
    (BookletPart == 1 & BlockInPart == 2 & currentindex == 16 & eventname == "Nav:GoTo") |
    (BookletPart == 2 & BlockInPart == 1 & currentindex == 2 & eventname == "Nav:GoTo")
  )

# Count occurrences for each different value of the column idstud
result <- filtered_data %>%
  group_by(idstud) %>%
  summarise(SQ01S03_goback = n())

# Merge the result back to the original data frame
result_goback_Q3 <- merge(result_Final_Answer_Q3, result, by = "idstud", all.x = TRUE)

result_goback_Q3$SQ01S03_goback[is.na(result_goback_Q3$SQ01S03_goback)] <- 0
```

```{r}
result_goback_Q3
# result_goback_Q3[c("idstud","eventname","currentindex","PSIVariableIdentifier","information","SQ01S03_goback")]
```


### Check how many times come back to the previous questions, including all the previous questions 
```{r}
# Filter the data based on the specified conditions
filtered_data <- result_time_complete_Q3 %>%
  filter(
    (BookletPart == 1 & BlockInPart == 2 & currentindex < 17 & currentindex>=15 & eventname == "Nav:GoTo") |
    (BookletPart == 2 & BlockInPart == 1 & currentindex < 3 & currentindex>=1 & eventname == "Nav:GoTo")
  )

# Count occurrences for each different value of the column idstud
result <- filtered_data %>%
  group_by(idstud) %>%
  summarise(SQ01S03_goback_all = n())

# Merge the result back to the original data frame
result_goback_all_Q3 <- merge(result_goback_Q3, result, by = "idstud", all.x = TRUE)

result_goback_all_Q3$SQ01S03_goback_all[is.na(result_goback_all_Q3$SQ01S03_goback_all)] <- 0
```

```{r}
result_goback_all_Q3
# result_goback_all_Q3[c("idstud","eventname","currentindex","PSIVariableIdentifier","information","SQ01S03_goback","SQ01S03_goback_all")]
```


### All Steps
Create a new dataframe by copy the original one.Create a new column step_str based on the other column.
If PSIVariableIdentifier == 5010 & information == 1, then step_str == "drop_coin"
If PSIVariableIdentifier == 5010 & information == 0, then step_str == "undo"
If PSIVariableIdentifier == 5011, then step_str == "input_" + value of the column information
If eventname == "Nav:GoTo" & new_currentindex == 3, then step_str == "Begin_Q3"
If eventname == "Nav:GoTo" & new_currentindex != 3, then step_str == "Nevigate_to_Q"+ value of the column new_currentindex
If eventname == "UI:Scrollbar", step_str == extract the value after event and scrollTop in the column information.
If eventname == "Btn:Calc", step_str == "Calculator" + extract the value after To in the column information.

Create a column new_currentindex by changing the column currentindex based on the rule below:
15 to 1, 16 to 2, 17 to 3, 18 to 4, 19 to 5, 20 to 6, 21 to 7, 22 to 8, 23 to 9, 24 to 10, 25 to 11, 26 to 12, 27 to 13, 28 to 14
```{r}
result_copy_Q3=data.frame(result_time_complete_Q3)
result_copy_Q3 <- result_copy_Q3 %>%
  mutate(new_currentindex = case_when(
    currentindex == 14 & BookletPart == 1 ~ 0,
    currentindex == 15 ~ 1,
    currentindex == 16 ~ 2,
    currentindex == 17 ~ 3,
    currentindex == 18 ~ 4,
    currentindex == 19 ~ 5,
    currentindex == 20 ~ 6,
    currentindex == 21 ~ 7,
    currentindex == 22 ~ 8,
    currentindex == 23 ~ 9,
    currentindex == 24 ~ 10,
    currentindex == 25 ~ 11,
    currentindex == 26 ~ 12,
    currentindex == 27 ~ 13,
    currentindex == 28 ~ 14,
    TRUE ~ currentindex  # Keep the original value if it doesn't match any rule
  ))
```

```{r}
result_str_Q3<- result_copy_Q3 %>%
  mutate(
    step_str = case_when(
      PSIVariableIdentifier == 5010 & information == 0 ~ "undo",
      PSIVariableIdentifier == 5010 & information == 1 ~ "drop_coin",
      PSIVariableIdentifier == 5011 & information != ""~ paste0("input_", information),
      PSIVariableIdentifier == 5011 & information == ""~ "input_NA",
      eventname == "Nav:GoTo" & new_currentindex == 3 ~ "Begin_Q3",
      eventname == "Nav:GoTo" & new_currentindex != 3 ~ paste0("Navigate_to_Q", new_currentindex),
      eventname == "UI:Scrollbar" ~ sub('.*"event":"([^"]+).*"scrollTop":(\\d+).*', '\\1 \\2', information),
      eventname == "Btn:Calc" ~ paste0("Calculator_", gsub('.*To":"(\\w+).*', '\\1', information)),
      TRUE ~ NA_character_  # If none of the conditions are met, set to NA
    )
  )
result_str_Q3[c("RowNumber","step_str","information")]
```


If the row of value of the column eventname == "UI:Scrollbar", then the value of the row of the column step_str substitute the white space to _.
```{r}
result_str_use_Q3 <- result_str_Q3 %>%
  mutate(step_str = ifelse(eventname == "UI:Scrollbar", gsub(" ", "_", step_str), step_str))
result_str_use_Q3[c("RowNumber","step_str","information")]
```

```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q3 %>%
  filter(!is.na(step_str)) %>%
  group_by(idstud) %>%
  mutate(SQ01S03_allsteps_list = list(step_str[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01S03_allsteps_seq <- sapply(df_filtered$SQ01S03_allsteps_list, function(x) paste0("(", toString(x), ")"))
```

```{r}
df_filtered <- select(df_filtered, -SQ01S03_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01S03_allsteps_seq")]

result_allsteps_Q3 <- merge(result_goback_all_Q3, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_allsteps_Q3=result_allsteps_Q3%>%
  group_by(idstud) %>%
  fill(SQ01S03_allsteps_seq, .direction = "downup")

result_allsteps_Q3
```

```{r}
result_allsteps_Q3[c("idstud","SQ01S03_allsteps_seq","information")]
```


### Timestamp for All Steps in milisecond
create a column to record the time taken for each single steps for each student to do this problem.
```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q3 %>%
  group_by(idstud) %>%
  mutate(SQ01S03_allsteps_list = list(SQ01S03_time[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01S03_allsteps_seq_time <- sapply(df_filtered$SQ01S03_allsteps_list, function(x) paste0("(", toString(x), ")"))

# filter the data and select column to merge
df_filtered <- select(df_filtered, -SQ01S03_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01S03_allsteps_seq_time")]

# merge the data
result_alls_time_Q3 <- merge(result_allsteps_Q3, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_alls_time_Q3=result_alls_time_Q3%>%
  group_by(idstud) %>%
  fill(SQ01S03_allsteps_seq_time, .direction = "downup")
result_alls_time_Q3
```


### Unit Time for All Steps
create a column to record the time taken for each single steps for each student to do this problem.
```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q3 %>%
  group_by(idstud) %>%
  mutate(SQ01S03_allsteps_list = list(SQ01S03_unit_time[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01S03_allsteps_timestamp <- sapply(df_filtered$SQ01S03_allsteps_list, function(x) paste0("(", toString(x), ")"))

# filter the data and select column to merge
df_filtered <- select(df_filtered, -SQ01S03_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01S03_allsteps_timestamp")]

# merge the data
result_all_unit_time_Q3 <- merge(result_alls_time_Q3, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_all_unit_time_Q3=result_all_unit_time_Q3%>%
  group_by(idstud) %>%
  fill(SQ01S03_allsteps_timestamp, .direction = "downup")
result_all_unit_time_Q3
```

### Total time taken
Create a new column SQ01S03_total_time. Group by idstud, SQ01S03_total_time = sum the value of the column SQ01S03_time for each idstud.

```{r}
# Group by idstud and calculate the sum of SQ01S03_time
result <- result_time_complete_Q3 %>%
  group_by(idstud) %>%
  summarise(SQ01S03_total_time = sum(SQ01S03_time, na.rm = TRUE))

# Merge the result back to the original data frame
result_total_time_Q3 <- merge(result_all_unit_time_Q3, result, by = "idstud", all.x = TRUE)
```

```{r}
result_total_time_Q3
# result_total_time_Q3[c("idstud","SQ01S03_time","SQ01S03_total_time")]
```


### Save the data
Filter the columns of data frame, only keep the new created columns and idstud, idstud_ori, ImportLogID, BlockName, BookletPart, BlockInPart.
```{r}
resultS03 <- result_total_time_Q3 %>%
  select(idstud, idstud_ori, ImportLogID, BlockName, BookletPart,BlockInPart, module_path, SQ01S03_undo, SQ01S03_drop_coin,SQ01S03_final_answer,SQ01S03_goback, SQ01S03_goback_all, SQ01S03_allsteps_seq, SQ01S03_allsteps_seq_time, SQ01S03_allsteps_timestamp, SQ01S03_total_time)
# resultS03
```


```{r}
resultS03_save = resultS03[!duplicated(resultS03), ]

condition1 <- resultS03_save$BlockName == 'SQ01' & 
              resultS03_save$BookletPart == 1 & 
              resultS03_save$BlockInPart == 2 

condition2 <- resultS03_save$BlockName == 'SQ01' & 
              resultS03_save$BookletPart == 2 & 
              resultS03_save$BlockInPart == 1

combined_condition <- condition1 | condition2

resultS03_save=resultS03_save%>%ungroup()%>% 
  filter(combined_condition)
resultS03_save
```

Save the dataframe
```{r}
# write_sav(resultS03_save, "../new_data/grade9/SQ12S03_PSI.sav")

# write.csv(resultS03_save, file = 'csv_SQ12S03_PSI.csv', row.names = FALSE)
```


