---
title: "S01"
author: "zifengxu"
date: "2024-01-28"
output: html_document
---

# Read the Data (PSI)
```{r}
library(haven)
library(dplyr)
library(stringr)
library(tidyr)
```

```{r}
df <- read_sav("../data/grade9/BSLNORZ7_PSI.sav")

# Explore the data
head(df)
```


Remove the rows of auto check for this question
```{r}
df_filter = df%>%
  filter(df$PSIVariableIdentifier != 5002 |is.na(df$PSIVariableIdentifier))
```

First of all, filter the rows and exactly one row above the row that satisfied one of the conditions below: 

Condition 1: BlockName=SQ01, BookletPart=1, BlockInPart=2, currentindex=15.

Condition 2: BlockName=SQ01, BookletPart=2, BlockInPart=1, currentindex=1.
```{r}
condition1 <- df_filter$BlockName == 'SQ01' & 
              df_filter$BookletPart == 1 & 
              df_filter$BlockInPart == 2 & 
              df_filter$currentindex == 15 

# Define the second condition
condition2 <- df_filter$BlockName == 'SQ01' & 
              df_filter$BookletPart == 2 & 
              df_filter$BlockInPart == 1 & 
              df_filter$currentindex == 1 

# Combine the two conditions using OR (|) as you want rows satisfying either condition
combined_condition <- condition1 | condition2

# Use filter to get the rows that satisfy the conditions and one row above
result1 <- df_filter %>% 
  filter(combined_condition | lag(combined_condition, default = FALSE))
```

```{r}
result1
```


### Module Path
Create a new column module_path based on (BlockName, BookletPart, BlockInPart, currentindex)
```{r}
result_use_Q1= transform(result1,
                       module_path = paste0("(", BlockName, ", ", BookletPart, ", ", BlockInPart, ")"))
result_use_Q1
```


### Time Taken
Filter the data that only related to students actions.
```{r}
result_filter_Q1=result_use_Q1 %>%filter(eventname %in% c("Nav:GoTo", "Response", "UI:Scrollbar", "Btn:Calc"))
result_filter_Q1
```

Create a new column SQ01S01_time based on two columns timeunixsec and timemilisec. The new column SQ01S01_time is created based on the rule below:

If currentindex == 15 & eventname == 'Nav:GoTo' or currentindex == 1 & eventname == 'Nav:GoTo', then return 0, otherwise, 

Let result_temp1 = the value of the row of the column timemilisec that satisfied the condition - the value of the previous row. 

If the result_temp1 is greater than 0, then result_temp2 = the value of the row of the column timeunixsec that satisfied the condition - the value of the previous row. Finally, the value of the row of the new created column SQ01S01_time = result_temp2+result_temp1/1000

If the result_temp1 is less than 0, then result_temp1 = 1000 + the value of the row of the column timemilisec that satisfied the condition. result_temp2 = the value of the row of the column timeunixsec that satisfied the condition - 1 - the value of the previous row. Finally, the value of the row of the new created column SQ01S01_time = result_temp2+result_temp1/1000
```{r}
result_time_complete_Q1=result_filter_Q1 %>%
  mutate(SQ01S01_time = ifelse(currentindex == 15 & eventname == 'Nav:GoTo' | currentindex == 1 & eventname == 'Nav:GoTo', 0, ifelse(timemilisec - lag(timemilisec) > 0,
                                      (timeunixsec - lag(timeunixsec) + (timemilisec - lag(timemilisec))/1000),
                                      (timeunixsec - lag(timeunixsec) - 1 + (timemilisec + 1000 - lag(timemilisec))/1000))))
# head(result_time_complete_Q1[c("idstud","module_path","SQ01S01_time")])

# result_time_complete_Q1
```

```{r}
result_time_complete_Q1 = result_time_complete_Q1%>%
  mutate(RowNumber = row_number())
```

```{r}
result_time_complete_Q1 
```


### Time in milisec
Create a new column SQ01S01_unit_time by merging the column timeunixsec and timemilisec together by following rule:
the new column SQ01S01_unit_time = timeunixsec * 1000 + timemilisec
```{r}
result_time_complete_Q1$SQ01S01_unit_time <- result_time_complete_Q1$timeunixsec * 1000 + result_time_complete_Q1$timemilisec

# show all the units
result_time_complete_Q1$SQ01S01_unit_time <- format(result_time_complete_Q1$SQ01S01_unit_time, scientific = FALSE)
result_time_complete_Q1
```


### Word Count 
Filter the data based on PSIVariableIdentifier == 5000 and IsFinalAnswer == 1.
```{r}
filtered_data <- result_time_complete_Q1 %>% filter(PSIVariableIdentifier == 5000)%>%
filter(IsFinalAnswer == 1)
```

```{r}
length(unique(filtered_data$idstud))
```

Create a new column SQ01S01_wordcount_punc that count the number of words in the column information including the punctuations.
```{r}
# Function to split words and punctuations
split_words_and_punctuations <- function(text) {
  # Match words and punctuations separately
  tokens <- strsplit(text, "(?<=[a-zA-Z0-9])(?=[[:punct:]])|(?<=[[:punct:]])(?=[a-zA-Z0-9])|\\s+", perl=TRUE)[[1]]
  # Remove empty tokens
  tokens <- tokens[tokens != ""]
  return(tokens)
}

# Apply the function to the information column
filtered_data$SQ01S01_wordcount_punc <- sapply(filtered_data$information, function(x) length(split_words_and_punctuations(x)))
```
Use (?<=[a-zA-Z0-9])(?=[[:punct:]])|(?<=[[:punct:]])(?=[a-zA-Z0-9]) split words and punctuations separately, and use |\\s+ matches one or more whitespace characters. Then count the total number of tokens.

excluding the punctuations
```{r}
filtered_data <- mutate(filtered_data, SQ01S01_wordcount_no_punc = str_count(gsub("[-[:punct:]]", " ", information), "\\S+"))
filtered_data[c("idstud","information","module_path","SQ01S01_wordcount_punc","SQ01S01_wordcount_no_punc")]
```

Merge the data to the original data
```{r}
# df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01S01_wordcount_punc","SQ01S01_wordcount_no_punc")]
df_filtered_merge = filtered_data[c("idstud","SQ01S01_wordcount_punc","SQ01S01_wordcount_no_punc")]
```

```{r}
# Merge the result back to the original data frame
result_wordcount_Q1 <- merge(result_time_complete_Q1, df_filtered_merge, by = "idstud", all.x = TRUE)
result_wordcount_Q1
```


### Raw Response
Create a new column SQ01S01_raw_response that is exactly same as the column information of the filtered data.
```{r}
filtered_data$SQ01S01_raw_response <- filtered_data$information

df_filtered_merge = filtered_data[c("idstud","SQ01S01_raw_response")]
```

```{r}
# Merge the result back to the original data frame
result_wordres_Q1 <- merge(result_wordcount_Q1, df_filtered_merge, by = "idstud", all.x = TRUE)
result_wordres_Q1
```


### All Steps
Create a new dataframe by copy the original one.Create a new column step_str based on the other column.
If PSIVariableIdentifier == 5000, then step_str == "write_answer"
If eventname == "Nav:GoTo" & new_currentindex == 1, then step_str == "Begin_Q1"
If eventname == "Nav:GoTo" & new_currentindex != 1, then step_str == "Nevigate_to_Q"+ value of the column new_currentindex
If eventname == "UI:Scrollbar", step_str == extract the value after event and scrollTop in the column information.
If eventname == "Btn:Calc", step_str == "Calculator" + extract the value after To in the column information.

Create a column new_currentindex by changing the column currentindex based on the rule below:
15 to 1, 16 to 2, 17 to 3, 18 to 4, 19 to 5, 20 to 6, 21 to 7, 22 to 8, 23 to 9, 24 to 10, 25 to 11, 26 to 12, 27 to 13, 28 to 14
```{r}
result_copy_Q1=data.frame(result_time_complete_Q1)
result_copy_Q1 <- result_copy_Q1 %>%
  mutate(new_currentindex = case_when(
    currentindex == 14 & BookletPart == 1 ~ 0,
    currentindex == 15 ~ 1,
    currentindex == 16 ~ 2,
    currentindex == 17 ~ 3,
    currentindex == 18 ~ 4,
    currentindex == 19 ~ 5,
    currentindex == 20 ~ 6,
    currentindex == 21 ~ 7,
    currentindex == 22 ~ 8,
    currentindex == 23 ~ 9,
    currentindex == 24 ~ 10,
    currentindex == 25 ~ 11,
    currentindex == 26 ~ 12,
    currentindex == 27 ~ 13,
    currentindex == 28 ~ 14,
    TRUE ~ currentindex  # Keep the original value if it doesn't match any rule
  ))
```

```{r}
result_str_Q1<- result_copy_Q1 %>%
  mutate(
    step_str = case_when(
      PSIVariableIdentifier == 5000 ~ "write_answer",
      eventname == "Nav:GoTo" & new_currentindex == 1 ~ "Begin_Q1",
      eventname == "Nav:GoTo" & new_currentindex != 1 ~ paste0("Navigate_to_Q", new_currentindex),
      eventname == "UI:Scrollbar" ~ sub('.*"event":"([^"]+).*"scrollTop":(\\d+).*', '\\1 \\2', information),
      eventname == "Btn:Calc" ~ paste0("Calculator_", gsub('.*To":"(\\w+).*', '\\1', information)),
      TRUE ~ NA_character_  # If none of the conditions are met, set to NA
    )
  )
result_str_Q1[c("RowNumber","step_str","information")]
```
 
If the row of value of the column eventname == "UI:Scrollbar", then the value of the row of the column step_str substitute the white space to _.
```{r}
result_str_use_Q1 <- result_str_Q1 %>%
  mutate(step_str = ifelse(eventname == "UI:Scrollbar", gsub(" ", "_", step_str), step_str))
result_str_use_Q1[c("RowNumber","step_str","information")]
```

```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q1 %>%
  filter(!is.na(step_str)) %>%
  group_by(idstud) %>%
  mutate(SQ01S01_allsteps_list = list(step_str[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01S01_allsteps_seq <- sapply(df_filtered$SQ01S01_allsteps_list, function(x) paste0("(", toString(x), ")"))
```

```{r}
df_filtered <- select(df_filtered, -SQ01S01_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01S01_allsteps_seq")]

result_allsteps_Q1 <- merge(result_wordres_Q1, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_allsteps_Q1=result_allsteps_Q1%>%
  group_by(idstud) %>%
  fill(SQ01S01_allsteps_seq, .direction = "downup")
# result_allsteps_Q1
```

```{r}
result_allsteps_Q1[c("idstud","SQ01S01_allsteps_seq","information")]
```


### Time Taken for All Steps
create a column to record the time taken for each single steps for each student to do this problem.
```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q1 %>%
  group_by(idstud) %>%
  mutate(SQ01S01_allsteps_list = list(SQ01S01_time[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01S01_allsteps_seq_time <- sapply(df_filtered$SQ01S01_allsteps_list, function(x) paste0("(", toString(x), ")"))

# filter the data and select column to merge
df_filtered <- select(df_filtered, -SQ01S01_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01S01_allsteps_seq_time")]

# merge the data
result_alls_time_Q1 <- merge(result_allsteps_Q1, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_alls_time_Q1=result_alls_time_Q1%>%
  group_by(idstud) %>%
  fill(SQ01S01_allsteps_seq_time, .direction = "downup")
result_alls_time_Q1
```


### Timestamp for All Steps in milisecond
create a column to record the time taken for each single steps for each student to do this problem.
```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q1 %>%
  group_by(idstud) %>%
  mutate(SQ01S01_allsteps_list = list(SQ01S01_unit_time[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01S01_allsteps_timestamp <- sapply(df_filtered$SQ01S01_allsteps_list, function(x) paste0("(", toString(x), ")"))

# filter the data and select column to merge
df_filtered <- select(df_filtered, -SQ01S01_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01S01_allsteps_timestamp")]

# merge the data
result_all_unit_time_Q1 <- merge(result_alls_time_Q1, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_all_unit_time_Q1=result_all_unit_time_Q1%>%
  group_by(idstud) %>%
  fill(SQ01S01_allsteps_timestamp, .direction = "downup")
result_all_unit_time_Q1
```

### Total time taken
Create a new column SQ01S01_total_time. Group by idstud, SQ01S01_total_time = sum the value of the column SQ01S01_time for each idstud.

```{r}
# Group by idstud and calculate the sum of SQ01S01_time
result <- result_time_complete_Q1 %>%
  group_by(idstud) %>%
  summarise(SQ01S01_total_time = sum(SQ01S01_time, na.rm = TRUE))

# Merge the result back to the original data frame
result_total_time_Q1 <- merge(result_all_unit_time_Q1, result, by = "idstud", all.x = TRUE)
```

```{r}
# result_total_time_Q1[c("idstud","SQ01S01_time","SQ01S01_total_time")]
result_total_time_Q1
```


### Save the data
Filter the columns of data frame, only keep the new created columns and idstud, idstud_ori, ImportLogID, BlockName, BookletPart, BlockInPart.
```{r}
resultS01 <- result_total_time_Q1 %>%
  select(idstud, idstud_ori, ImportLogID, BlockName, BookletPart,BlockInPart, module_path, SQ01S01_wordcount_punc, SQ01S01_wordcount_no_punc, SQ01S01_raw_response, SQ01S01_allsteps_seq, SQ01S01_allsteps_seq_time, SQ01S01_allsteps_timestamp, SQ01S01_total_time)
# resultS01
```


```{r}
resultS01_save = resultS01[!duplicated(resultS01), ]

condition1 <- resultS01_save$BlockName == 'SQ01' & 
              resultS01_save$BookletPart == 1 & 
              resultS01_save$BlockInPart == 2 

condition2 <- resultS01_save$BlockName == 'SQ01' & 
              resultS01_save$BookletPart == 2 & 
              resultS01_save$BlockInPart == 1

combined_condition <- condition1 | condition2

resultS01_save=resultS01_save%>%ungroup()%>% 
  filter(combined_condition)
resultS01_save
```

Save the dataframe
```{r}
# write_sav(resultS01_save, "../new_data/grade9/SQ12S01_PSI.sav")

# write.csv(resultS01_save, file = 'csv_SQ12S01_PSI.csv', row.names = FALSE)
```


