---
title: "S04"
author: "zifengxu"
date: "2023-12-12"
output: html_document
---

# Read the Data (PSI)
```{r}
library(haven)
library(dplyr)
library(stringr)
library(tidyr)
library(jsonlite)
```

```{r}
df <- read_sav("../data/grade9/BSLNORZ7_PSI.sav")

# Explore the data
head(df)
```

Remove the rows of auto check for this question
```{r}
df_filter = df%>%
  filter(df$PSIVariableIdentifier != 5016 |is.na(df$PSIVariableIdentifier))
```

Fiter the rows and exactly one row above the row that satisfied one of the conditions below: 

Condition 1: BlockName=SQ01, BookletPart=1, BlockInPart=2, currentindex=20.

Condition 2: BlockName=SQ01, BookletPart=2, BlockInPart=1, currentindex=6.

```{r}
condition1 <- df_filter$BlockName == 'SQ01' & 
              df_filter$BookletPart == 1 & 
              df_filter$BlockInPart == 2 & 
              df_filter$currentindex == 18

# Define the second condition
condition2 <- df_filter$BlockName == 'SQ01' & 
              df_filter$BookletPart == 2 & 
              df_filter$BlockInPart == 1 & 
              df_filter$currentindex == 4 

# Combine the two conditions using OR (|) as you want rows satisfying either condition
combined_condition <- condition1 | condition2

# Use filter to get the rows that satisfy the conditions and one row below
result4 <- df_filter %>% 
  filter(combined_condition | lag(combined_condition, default = FALSE))

result4
```


### Module Path
Create a new column module_path based on (BlockName, BookletPart, BlockInPart, currentindex)
```{r}
result_use_Q4= transform(result4,
                       module_path = paste0("(", BlockName, ", ", BookletPart, ", ", BlockInPart, ")"))
head(result_use_Q4)
```

```{r}
unique(result_use_Q4$eventname)
```


### Time
```{r}
result_filter_Q4=result_use_Q4 %>%filter(eventname %in% c("Nav:GoTo", "Response", "UI:Scrollbar", "Btn:Calc"))
result_filter_Q4
```


```{r}
result_time_complete_Q4=result_filter_Q4 %>%
  mutate(SQ01S04_time = ifelse(currentindex == 18 & eventname == 'Nav:GoTo' | currentindex == 4 & eventname == 'Nav:GoTo', 0, ifelse(timemilisec - lag(timemilisec) > 0,
                                      (timeunixsec - lag(timeunixsec) + (timemilisec - lag(timemilisec))/1000),
                                      (timeunixsec - lag(timeunixsec) - 1 + (timemilisec + 1000 - lag(timemilisec))/1000))))
# head(result_time_complete_Q4[c("idstud","module_path","SQ01S04_time")])

# result_time_complete_Q4
```

```{r}
result_time_complete_Q4 = result_time_complete_Q4%>%
  mutate(RowNumber = row_number())
```

```{r}
result_time_complete_Q4$SQ01S04_unit_time <- result_time_complete_Q4$timeunixsec * 1000 + result_time_complete_Q4$timemilisec

# show all the units
result_time_complete_Q4$SQ01S04_unit_time <- format(result_time_complete_Q4$SQ01S04_unit_time, scientific = FALSE)
```

```{r}
result_time_complete_Q4
```


### Fill in declicking 
If the column PSIVariableIdentifier == 5014 and the column information is empty, then fill "declick" into the column information.
```{r}
result_declick_Q4 <- result_time_complete_Q4 %>%
  mutate(information = ifelse(PSIVariableIdentifier == 5014 & information == "", "declick", information))

result_declick_Q4
```


### Check how many times come back to the previous question.
```{r}
result_time_complete_Q4[c("currentindex","eventname","information")]
```

Create a new column SQ01S04_goback. Filter the data according to BookletPart=1, BlockInPart=2, currentindex=17, eventname="Nav:GoTo" or BookletPart=2, BlockInPart=1, currentindex=3, eventname="Nav:GoTo". Count how the occurance for each different value of the column idstud.

```{r}
# Filter the data based on the specified conditions
filtered_data <- result_time_complete_Q4 %>%
  filter(
    (BookletPart == 1 & BlockInPart == 2 & currentindex == 17 & eventname == "Nav:GoTo") |
    (BookletPart == 2 & BlockInPart == 1 & currentindex == 3 & eventname == "Nav:GoTo")
  )

# Count occurrences for each different value of the column idstud
result <- filtered_data %>%
  group_by(idstud) %>%
  summarise(SQ01S04_goback = n())

# Merge the result back to the original data frame
result_goback_Q4 <- merge(result_time_complete_Q4, result, by = "idstud", all.x = TRUE)

result_goback_Q4$SQ01S04_goback[is.na(result_goback_Q4$SQ01S04_goback)] <- 0
```

```{r}
result_goback_Q4
# result_goback_Q4[c("idstud","eventname","currentindex","PSIVariableIdentifier","information","SQ01S04_goback")]
```


### Check how many times come back to the previous questions, including all the previous questions 
```{r}
# Filter the data based on the specified conditions
filtered_data <- result_time_complete_Q4 %>%
  filter(
    (BookletPart == 1 & BlockInPart == 2 & currentindex < 18 & currentindex>=15 & eventname == "Nav:GoTo") |
    (BookletPart == 2 & BlockInPart == 1 & currentindex < 4 & currentindex>=1 & eventname == "Nav:GoTo")
  )

# Count occurrences for each different value of the column idstud
result <- filtered_data %>%
  group_by(idstud) %>%
  summarise(SQ01S04_goback_all = n())

# Merge the result back to the original data frame
result_goback_all_Q4 <- merge(result_goback_Q4, result, by = "idstud", all.x = TRUE)

result_goback_all_Q4$SQ01S04_goback_all[is.na(result_goback_all_Q4$SQ01S04_goback_all)] <- 0
```

```{r}
result_goback_all_Q4
# result_goback_all_Q4[c("idstud","eventname","currentindex","PSIVariableIdentifier","information","SQ01S04_goback","SQ01S04_goback_all")]
```


### rewrite(write) the answer?
```{r}
# df_use = result_time_complete_Q4%>%
#   mutate(RowNumber = row_number())
# df_use
df_use = data.frame(result_time_complete_Q4)
```

Create a new column ans_change where all the values are 0 at first. Go through the rows, if the column eventname == "Nav:GoTo" and BlockInPart == 1 and currentindex == 4 or eventname == "Nav:GoTo" and BlockInPart == 2 and currentindex == 18, until the next time there exist rows satisfied this condition, if there exists rows that satisfied the column eventname == "Response" and PSIVariableIdentifier== 5014 and information is not NA, then the new column ans_change of the row satisfied this condition = 1, otherwise 0.
```{r}
library(dplyr)

# Assuming df_use is your original data frame
df_use <- df_use %>%
  mutate(ans_change = 0)  # Create a new column initialized with 0

# Iterate through the rows
for (i in 1:(nrow(df_use) - 1)) {
  if ((df_use$eventname[i] == "Nav:GoTo" && df_use$BlockInPart[i] == 1 && df_use$currentindex[i] == 4) ||
      (df_use$eventname[i] == "Nav:GoTo" && df_use$BlockInPart[i] == 2 && df_use$currentindex[i] == 18)) {
    
    # Check for the specified condition in subsequent rows
    j <- i + 1
    while (j <= nrow(df_use) && !(df_use$eventname[j] == "Nav:GoTo" && (df_use$currentindex[j] == 4 || df_use$currentindex[j] == 18))) {
      if (df_use$eventname[j] == "Response" && df_use$PSIVariableIdentifier[j] == 5014 && !is.na(df_use$information[j])) {
        df_use$ans_change[i] <- 1
        break
      }
      j <- j + 1
    }
  }
}
```

```{r}
df_use[c("eventname","currentindex","information","PSIVariableIdentifier","SQ01S04_time","ans_change","RowNumber")]
```

Filter the data based on the column eventname == "Nav:GoTo" and currentindex == 18 or eventname == "Nav:GoTo" and currentindex == 4.Then group by the student id, the new column SQ01S04_answer_rewrite should created based on the sequence of the value of the column information in the order of if the row number is higher, their sequence should be higher.
```{r}
df_filtered <- df_use %>%
  filter((eventname == "Nav:GoTo" & currentindex == 18) |
         (eventname == "Nav:GoTo" & currentindex == 4))%>%
  group_by(idstud) %>%
  mutate(SQ01S04_sequence_list = list(ans_change[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01S04_answer_rewrite <- sapply(df_filtered$SQ01S04_sequence_list, function(x) paste0("(", toString(x), ")"))

# df_filtered
```

```{r}
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01S04_answer_rewrite")]
```

```{r}
result_answer_rewrite_Q4 <- merge(result_goback_all_Q4, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)
# result_answer_rewrite_Q4
```

Then fill in all the missing value of the column SQ01S04_answer_rewrite according to the student name.
```{r}
result_answer_rewrite_Q4=result_answer_rewrite_Q4%>%
  group_by(idstud) %>%
  fill(SQ01S04_answer_rewrite, .direction = "downup")
result_answer_rewrite_Q4
```


### Answer in Sequence
Filter the data by PSIVariableIdentifier == 5014. Then group by the student id, the new column SQ01S04_answer_sequence should created based on the sequence of the value of the column information in the order of if the row number is higher, their sequence should be higher.

```{r}
# Filter data and arrange by idstud and information
df_filtered <- result_time_complete_Q4 %>%
  filter(PSIVariableIdentifier == 5014)%>%
  group_by(idstud) %>%
  mutate(SQ01S04_sequence_list = list(information[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01S04_answer_sequence <- sapply(df_filtered$SQ01S04_sequence_list, function(x) paste0("(", toString(x), ")"))

# df_filtered
```

Create a new dataframe by merge the two dataframes by using the column idstud, PSIVariableIdentifier, and RowNumber. Since each student has unique student id and response id to answer different questions.
```{r}
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01S04_answer_sequence")]
```

```{r}
# 7,209 rows
result_ans_seq_Q4 <- merge(result_answer_rewrite_Q4, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)
# result_ans_seq_Q4
```

Then fill in all the missing value of the column SQ01S02_answer_sequence according to the student name.
```{r}
result_ans_seq_Q4=result_ans_seq_Q4%>%
  group_by(idstud) %>%
  fill(SQ01S04_answer_sequence, .direction = "downup")
result_ans_seq_Q4
```


### Final answer
Filter the column by taking PSIVariableIdentifier == 5014 and IsFinalAnswer == 1. Then create a new column named SQ01S04_final_answer and this column is exactly the same with the column information.
```{r}
# Filter data and arrange by idstud 
result_final_Q4=result_time_complete_Q4 %>%
  filter(PSIVariableIdentifier == 5014 & IsFinalAnswer == 1)%>%
  mutate(SQ01S04_final_answer = information)

# merge the dataframe
df_filtered_merge = result_final_Q4[c("idstud","RowNumber","SQ01S04_final_answer")]
result_final_result_Q4 <- merge(result_ans_seq_Q4, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

# fill the missing value
result_final_result_Q4=result_final_result_Q4%>%
  group_by(idstud) %>%
  fill(SQ01S04_final_answer, .direction = "downup")

result_final_result_Q4
```


### Check the total time taken for each time

```{r}
df_test = data.frame(result_time_complete_Q4)
# df_test
```

Go through all the row of the data by sequence of the row. Create a new column time_sum based on the column SQ01S04_time. If the SQ01S04_time is NA, then the new column time_sum = 0, otherwise the new column time_sum is calculated by keep adding the value of the column SQ01S04_time, until the value of the column SQ01S04_time is NA. Then the new_column time_sum = 0, and then add the value of the column SQ01S04_time again.

```{r}
# Create a new column time_sum
df_test$time_sum <- 0

# Iterate through the rows
for (i in 2:nrow(df_test)) {
  # If SQ01S04_time is NA, set time_sum to 0
  if (df_test$SQ01S04_time[i] == 0) {
    df_test$time_sum[i] <- 0
  } else {
    # Add the current SQ01S04_time to time_sum
    df_test$time_sum[i] <- df_test$time_sum[i - 1] + df_test$SQ01S04_time[i]
  }
}

# Print the result
# print(df_test)
```

```{r}
condition <- df_test$SQ01S04_time == 0
df_test_filter = df_test%>% 
  filter(lead(condition, default = FALSE))
# df_test_filter
```

```{r}
# Filter data and arrange by idstud and information
df_filtered <- df_test_filter %>%
  group_by(idstud) %>%
  mutate(SQ01S04_sequence_list = list(time_sum[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01S04_seq_time <- sapply(df_filtered$SQ01S04_sequence_list, function(x) paste0("(", toString(x), ")"))

# filter the merge data
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01S04_seq_time")]

result_time_seq_Q4 <- merge(result_final_result_Q4, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_time_seq_Q4=result_time_seq_Q4%>%
  group_by(idstud) %>%
  fill(SQ01S04_seq_time, .direction = "downup")
```

```{r}
result_time_seq_Q4
```


### All Steps
Create a new dataframe by copy the original one.Create a new column step_str based on the other column.
If PSIVariableIdentifier == 5014, then step_str == "select_option_" + value of the column information
If eventname == "Nav:GoTo" & new_currentindex == 4, then step_str == "Begin_Q4"
If eventname == "Nav:GoTo" & new_currentindex != 4, then step_str == "Nevigate_to_Q"+ value of the column new_currentindex
If eventname == "UI:Scrollbar", step_str == extract the value after event and scrollTop in the column information.
If eventname == "Btn:Calc", step_str == "Calculator" + extract the value after To in the column information.

Create a column new_currentindex by changing the column currentindex based on the rule below:
15 to 1, 16 to 2, 17 to 3, 18 to 4, 19 to 5, 20 to 6, 21 to 7, 22 to 8, 23 to 9, 24 to 10, 25 to 11, 26 to 12, 27 to 13, 28 to 14
```{r}
result_copy_Q4=data.frame(result_declick_Q4)
result_copy_Q4 <- result_copy_Q4 %>%
  mutate(new_currentindex = case_when(
    currentindex == 14 & BookletPart == 1 ~ 0,
    currentindex == 15 ~ 1,
    currentindex == 16 ~ 2,
    currentindex == 17 ~ 3,
    currentindex == 18 ~ 4,
    currentindex == 19 ~ 5,
    currentindex == 20 ~ 6,
    currentindex == 21 ~ 7,
    currentindex == 22 ~ 8,
    currentindex == 23 ~ 9,
    currentindex == 24 ~ 10,
    currentindex == 25 ~ 11,
    currentindex == 26 ~ 12,
    currentindex == 27 ~ 13,
    currentindex == 28 ~ 14,
    TRUE ~ currentindex  # Keep the original value if it doesn't match any rule
  ))
```

```{r}
result_str_Q4<- result_copy_Q4 %>%
  mutate(
    step_str = case_when(
      PSIVariableIdentifier == 5014 ~ paste0("select_option_", information),
      eventname == "Nav:GoTo" & new_currentindex == 4 ~ "Begin_Q4",
      eventname == "Nav:GoTo" & new_currentindex != 4 ~ paste0("Navigate_to_Q", new_currentindex),
      eventname == "UI:Scrollbar" ~ sub('.*"event":"([^"]+).*"scrollTop":(\\d+).*', '\\1 \\2', information),
      eventname == "Btn:Calc" ~ paste0("Calculator_", gsub('.*To":"(\\w+).*', '\\1', information)),
      TRUE ~ NA_character_  # If none of the conditions are met, set to NA
    )
  )
result_str_Q4[c("RowNumber","step_str","information")]
```
 
If the row of value of the column eventname == "UI:Scrollbar", then the value of the row of the column step_str substitute the white space to _.
```{r}
result_str_use_Q4 <- result_str_Q4 %>%
  mutate(step_str = ifelse(eventname == "UI:Scrollbar", gsub(" ", "_", step_str), step_str))
result_str_use_Q4[c("RowNumber","step_str","information")]
```

```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q4 %>%
  filter(!is.na(step_str)) %>%
  group_by(idstud) %>%
  mutate(SQ01S04_allsteps_list = list(step_str[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01S04_allsteps_seq <- sapply(df_filtered$SQ01S04_allsteps_list, function(x) paste0("(", toString(x), ")"))
```

```{r}
df_filtered <- select(df_filtered, -SQ01S04_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01S04_allsteps_seq")]

result_allsteps_Q4 <- merge(result_time_seq_Q4, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_allsteps_Q4=result_allsteps_Q4%>%
  group_by(idstud) %>%
  fill(SQ01S04_allsteps_seq, .direction = "downup")

result_allsteps_Q4
```

```{r}
result_allsteps_Q4[c("idstud","SQ01S04_allsteps_seq","information")]
```


### Timestamp for All Steps in milisecond
create a column to record the time taken for each single steps for each student to do this problem.
```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q4 %>%
  group_by(idstud) %>%
  mutate(SQ01S04_allsteps_list = list(SQ01S04_time[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01S04_allsteps_seq_time <- sapply(df_filtered$SQ01S04_allsteps_list, function(x) paste0("(", toString(x), ")"))

# filter the data and select column to merge
df_filtered <- select(df_filtered, -SQ01S04_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01S04_allsteps_seq_time")]

# merge the data
result_alls_time_Q4 <- merge(result_allsteps_Q4, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_alls_time_Q4=result_alls_time_Q4%>%
  group_by(idstud) %>%
  fill(SQ01S04_allsteps_seq_time, .direction = "downup")
result_alls_time_Q4
```


### Unit Time for All Steps
create a column to record the time taken for each single steps for each student to do this problem.
```{r}
# Filter data and create a list to record all the steps needed.
df_filtered <- result_str_use_Q4 %>%
  group_by(idstud) %>%
  mutate(SQ01S04_allsteps_list = list(SQ01S04_unit_time[order(row_number())]))

# Convert the list column to a new column with the desired format, and then drop the list column
df_filtered$SQ01S04_allsteps_timestamp <- sapply(df_filtered$SQ01S04_allsteps_list, function(x) paste0("(", toString(x), ")"))

# filter the data and select column to merge
df_filtered <- select(df_filtered, -SQ01S04_allsteps_list)
df_filtered_merge = df_filtered[c("idstud","RowNumber","SQ01S04_allsteps_timestamp")]

# merge the data
result_all_unit_time_Q4 <- merge(result_alls_time_Q4, df_filtered_merge, by = c("idstud","RowNumber"), all = TRUE)

result_all_unit_time_Q4=result_all_unit_time_Q4%>%
  group_by(idstud) %>%
  fill(SQ01S04_allsteps_timestamp, .direction = "downup")
result_all_unit_time_Q4
```

### Total time taken
Create a new column SQ01S04_total_time. Group by idstud, SQ01S04_total_time = sum the value of the column SQ01S04_time for each idstud.

```{r}
# Group by idstud and calculate the sum of SQ01S04_time
result <- result_time_complete_Q4 %>%
  group_by(idstud) %>%
  summarise(SQ01S04_total_time = sum(SQ01S04_time, na.rm = TRUE))

# Merge the result back to the original data frame
result_total_time_Q4 <- merge(result_all_unit_time_Q4, result, by = "idstud", all.x = TRUE)
```

```{r}
result_total_time_Q4
# result_total_time_Q4[c("idstud","SQ01S04_time","SQ01S04_total_time")]
```


### Save the data
Filter the columns of data frame, only keep the new created columns and idstud, idstud_ori, ImportLogID, BlockName, BookletPart, BlockInPart.
```{r}
resultS04 <- result_total_time_Q4 %>%
  select(idstud, idstud_ori, ImportLogID, BlockName, BookletPart,BlockInPart, module_path, SQ01S04_goback, SQ01S04_goback_all, SQ01S04_answer_rewrite, SQ01S04_answer_sequence, SQ01S04_final_answer, SQ01S04_seq_time, SQ01S04_allsteps_seq, SQ01S04_allsteps_seq_time, SQ01S04_allsteps_timestamp, SQ01S04_total_time)
# resultS04
```


```{r}
resultS04_save = resultS04[!duplicated(resultS04), ]

condition1 <- resultS04_save$BlockName == 'SQ01' & 
              resultS04_save$BookletPart == 1 & 
              resultS04_save$BlockInPart == 2 

condition2 <- resultS04_save$BlockName == 'SQ01' & 
              resultS04_save$BookletPart == 2 & 
              resultS04_save$BlockInPart == 1

combined_condition <- condition1 | condition2

resultS04_save=resultS04_save%>%ungroup()%>% 
  filter(combined_condition)
resultS04_save
```

Save the dataframe
```{r}
# write_sav(resultS04_save, "../new_data/grade9/SQ12S04_PSI.sav")

# write.csv(resultS04_save, file = 'csv_SQ12S04_PSI.csv', row.names = FALSE)
```



